---
title: Exercise 8.8
date: "March 2024"
output: html_document 
author: Kári Hlynsson
---

In the lab, a classificiation tree was applied to the `Carseats`
data after converting `Sales` into a qualitative reponse variable.
Now we will seek to predict `Sales` using regression trees and related
approaches, treating the response as a quantitative variable.

## Part (a) 
Split the data set into a training set and a test set.

```{r}
library(ISLR2)
data(Carseats)

set.seed(1)

n <- nrow(Carseats)
p <- ncol(Carseats) - 1
train_prop <- 0.75

s <- sample(
  c(TRUE, FALSE),
  size = n,
  replace = TRUE,
  prob = c(train_prop, 1 - train_prop)
)

train <- Carseats[s, ]
test <- Carseats[!s, ]
```

## Part (b)
Fit a regression tree to the training set. Plot the tree, and intrepret
the results. What test MSE do you obtain?

Start by fitting the tree:

```{r}
library(tree)

tree_cars <- tree(Sales ~ ., data = train)
summary(tree_cars)
```

The predictors in the data used in the regression tree were `ShelveLoc`,
`Price`, `CompPrice`, `Age`, `Income` and `Advertising`, i.e. 6 out of 11
predictors.

Produce the requested plot:

```{r}
plot(tree_cars)
text(tree_cars, pretty = 0)
```

Now let's compute the test MSE:

```{r}
tree_pred <- predict(tree_cars, test)
tree_mse <- mean((tree_pred - test$Sales)^2)
```

The test MSE is `r tree_mse`.

## Part (c)
Use cross-validation in order to determine the optimal level of tree
complexity. Does pruning the tree improve the test MSE?

```{r}
cv_cars <- cv.tree(tree_cars)

par(mfrow = c(1, 2))
plot(
  cv_cars$size,
  cv_cars$dev,
  type = "b",
  xlab = "No. Terminal Nodes",
  ylab = "Deviance"
)

plot(
  cv_cars$k,
  cv_cars$dev,
  type = "b",
  xlab = "k",
  ylab = "Deviance"
)
```

The deviance of the regression tree is highest for \(n = 1\) terminal nodes 
and decreases up to \(n = 5\) whereupon it starts fluctuating a bit up and 
down. The value of \(n\) which minimises the deviance is \(n = 9\) as can be 
seen by the code below. Thus the optimal value of terminal nodes seems to be
nine nodes.

```{r}
n_opt <- cv_cars$size[which.min(cv_cars$dev)]
print(n_opt)
```

Let's try pruning the regression tree `tree_cars` to yield an optimal tree
with nine terminal nodes:

```{r}
cars_prune <- prune.tree(tree_cars, best = n_opt)

plot(cars_prune)
text(cars_prune, pretty = 0)
```
Measure the MSE:

```{r}
prune_pred <- predict(cars_prune, test)
prune_mse <- mean((prune_pred - test$Sales)^2)
```

The MSE of the pruned tree with the optimal value of \(n = 9\) is `r opt_mse`,
which is slightly higher than that of the original model where all the 
predictors were involved. However, this tree is a bit smaller and thus 
slightly more interpretable.

## Part (d)
Use the bagging approach in order to analyse this data. What test MSE do you
obtain? Use the `importance()` function to determine which variables are most
important.

```{r}
library(randomForest)

cars_bagging <- randomForest(Sales ~ ., data = train, mtry = p - 1)
bag_pred <- predict(cars_bagging, test)
bag_mse <- mean((bag_pred - test$Sales)^2)
```

The bagging MSE is `r bag_mse`, which is a drastic improvement on the pruned
regression tree.

```{r}
bagging_importance <- importance(cars_bagging)
ord <- order(bagging_importance, decreasing = TRUE)

print(bagging_importance[ord, , drop = FALSE] |> head(5))
varImpPlot(cars_bagging)
```

The top 5 most important variables are `ShelveLoc`, `Price`, `CompPrice`,
`Age` and `Advertising`.

## Part (e)
Use random forests to analyse this data. What test MSE do you obtain? Use the 
`importance()` function to determine which variables are most important.
Describe the effect of \(m\), the number of variables considered at each split,
on the error rate obtained.

Fit a random forest model with \(m = p/3\):

```{r}
cars_rf <- randomForest(Sales ~ ., data = train)

rf_pred <- predict(cars_rf, test)
rf_mse <- mean((rf_pred - test$Sales)^2)
```

The MSE is `r rf_mse`, which is slightly higher than that of the bagging
model. Variable importances:

```{r}
rf_importance <- importance(cars_rf)
ord <- order(rf_importance, decreasing = TRUE)

print(rf_importance[ord, , drop = FALSE] |> head(5))
varImpPlot(cars_rf)
```

The top 5 most important variables have remained unchanged, although `Age` has
overtaken `CompPrice` in terms of variable importance.

Let's try looking at how the error rate changes with \(m\):

```{r}
errs <- sapply(1:p, function(i) {
  cars_rf_i <- randomForest(Sales ~ ., data = train)
  rf_i_pred <- predict(cars_rf_i, test)
  rf_mse <- mean((rf_i_pred - test$Sales)^2)
})

plot(1:p, errs, type = "b", xlab = "m", ylab = "Test MSE")
```

The test MSE seems to be relatively stable for different values for \(m\) but
is optimal for \(m = 5\), with the test MSE being `r errs[5]`. This is
considerably higher than \(\sqrt{p}\) or \(p/3\) as per the gold standard, but
is consistent with our findings that the bagging model performed better. This
may be a better parameter value for \(m\).

## Part (f)
Now analyse the data using BART, and report your results.

```{r}
library(BART)

x <- Carseats[, 2:(p + 1)]
y <- Carseats[, 1]

x_train <- x[s, ]
x_test <- x[!s, ]
y_train <- y[s]
y_test <- y[!s]
```

Compute the test MSE and look at variable importances in terms of occurence
in the ensemble:

```{r}
cars_bart <- gbart(x_train, y_train, x.test = x_test)

bart_pred <- cars_bart$yhat.test.mean
bart_mse <- mean((bart_pred - test$Sales)^2)
```

The BART MSE is `r bart_mse`, the best so far.

```{r}
mses <- c(tree_mse, prune_mse, bag_mse, rf_mse, errs[5], bart_mse)
names <- c(
  "Tree", "Pruned Tree", "Bagging", "Random Forest",
  "CV Random Forest", "BART"
)

barplot(
  mses,
  names.arg = names,
  xlab = "Method",
  ylab = "Test MSE",
  cex.names = 0.6
)
```
