---
title: Exercise 8.9
date: March 2024
output: html_document
author: Kári Hlynsson
---

This problem involes the `OJ` data set which is part of the `ISLR2` package.

## Part (a)

Create a training set containing a random sample of 800 observations, and a 
test set containing the remaining observations.

```{r}
library(ISLR2)
data(OJ)

set.seed(1)

n <- nrow(OJ)
p <- ncol(OJ)

s <- sample(1:n, 800)
r <- 1:n %in% s

train <- OJ[r, ]
test <- OJ[!r, ]
```

## Part (b)
Fit a tree to the training data, with `Purchase` as the response and the other
variables as predictors. Use the `summary()` function to produce summary
statistics about the tree, and describe the results obtained. What is the
training error rate? How many terminal nodes does the tree have?

```{r}
library(tree)

purchase_tree <- tree(Purchase ~ ., data = train)
summary(purchase_tree)
```

The number of terminal nodes is 8 as seen in the summary output. The training
error rate is in terms of misclassification and is \(133 / 800 = 0.1662\).

## Part (c)
Type in the name of the tree object in order to get a detailed text output.
Pick one of the terminal nodes, and interpret the information displayed.

```{r}
purchase_tree
```

Node 24 states that if `PctDiscMM < 0.196196`, classify the observation as 
`CH`, whereas node 25 states if `PctDiscMM > 0.196196`, classify it as `MM`.

## Part (d)
Create a plot of the tree, and interpret the results.

```{r}
plot(purchase_tree)
text(purchase_tree, pretty = 0)
```

## Part (e)
Predict the reponse on the test data, and produce a confusion matrix comparing
the test labels to the predicted test labels. What is the test error rate?

```{r}
library(caret)

tree_pred <- predict(purchase_tree, test, type = "class")
tree_conf <- confusionMatrix(tree_pred, test$Purchase)
tree_conf
```

The test error rate is \(1 - 0.8296 = 0.1704\). The sensitivity is \(0.9524\) 
and the specificity is \(0.6275\), so the classifier isn't all that bad.

## Part (f)
Apply the `cv.tree()` function to the training set in order to determine the
optimal tree size.

```{r}
purchase_cv <- cv.tree(purchase_tree, FUN = prune.misclass)

par(mfrow = c(1, 2))

plot(
  purchase_cv$size,
  purchase_cv$dev,
  type = "b",
  xlab = "Size",
  ylab = "Deviance"
)

plot(
  purchase_cv$k,
  purchase_cv$dev,
  type = "b",
  xlab = "k",
  ylab = "Deviance"
)
```

```{r}
n_opt <- purchase_cv$size[which.min(purchase_cv$dev)]
```

The optimal number of terminal nodes is `r n_opt`.

## Part (g)
Produce a plot with tree size on the \(x\)-axis and cross-validated
classification error rate on the \(y\)-axis.

```{r}
plot(purchase_cv$size, purchase_cv$dev,
  xlab = "Terminal node size",
  ylab = "CV classifiction error rate", type = "b"
)
```

## Part (h)
Which tree size corresponds to the lowest cross-validated classification
error rate?

`r n_opt`.

## Part (i)
Produce a pruned tree corresponding to the optimal tree size obtained using
cross-validation. If cross-validation does not lead to selection of a pruned
tree, then create a pruned tree with five terminal nodes.

```{r}
purchase_prune <- prune.tree(purchase_tree, best = n_opt)

prune_pred <- predict(purchase_prune, test, type = "class")
prune_conf <- confusionMatrix(prune_pred, test$Purchase)
prune_conf
```

## Part (j)
Compare the training error rates between the pruned and unpruned trees. Which
is higher?

The training error rate is slightly lower for the larger tree with all of the
predictors (`r summary(purchase_tree)$misclass[1] / summary(purchase_tree)$
misclass[2]`) when compared to the pruned tree  (`r summary(purchase_prune)$
misclass[1] / summary(purchase_prune)$misclass[2]`).

## Part (k)
Compare the test error rates between the pruned and unpruned trees. Which is
higher?

The test error rate for the original tree is 
`r 1 - tree_conf$overall["Accuracy"]` whereas for the pruned tree it is 
`r 1 - prune_conf$overall["Accuracy"]`. This is consistent with the theory.
