---
title: Exercise 8.12
date: March 2024
output: html_document 
author: Kári Hlynsson
---

```{r setup, include=FALSE, message = FALSE}
panel_hist <- function(x, ...) {
  usr <- par("usr")
  par(usr = c(usr[1:2], 0, 1.5))
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks
  nb <- length(breaks)
  y <- h$counts
  y <- y / max(y)
  rect(breaks[-nb], 0, breaks[-1], y, ...)
}

panel_cor <- function(x, y, digits = 2, prefix = "", cex_cor, ...) {
  par(usr = c(0, 1, 0, 1))
  r <- cor(x, y)
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste0(prefix, txt)
  text(0.5, 0.5, txt, cex = 1.5)
}

pair_plot <- function(x) {
  pairs(x, diag.panel = panel_hist, upper.panel = panel_cor)
}
```

Apply boosting, bagging, random forests, and BART to a data set of your choice.
Be sure to fit the models on a training set and to evaluate their performance
on a test set. How accurate are the results compared to simple methods like
linear or logistic regression? Which of these approaches yields the best
performance?

```{r}
library(ISLR2)
library(Biostatistics)

set.seed(1)

data(longevity)

n <- nrow(longevity)
p <- ncol(longevity)

# convert some stuff to factors
longevity$class <- as.factor(longevity$class)
longevity$volancy <- as.factor(longevity$volancy)
longevity$fossoriallity <- as.factor(longevity$fossoriallity)
longevity$foraging_environment <- as.factor(longevity$foraging_environment)
longevity$daily_activity <- as.factor(longevity$daily_activity)

longevity <- subset(longevity, select = -c(species, order))
```

Pair plot of the data:

```{r}
pair_plot(longevity)
```

Let's log-transform the mass, since it seems to be exponentially distributed:

```{r}
longevity$log_mass_g <- log(longevity$mass_g, base = 10)
longevity <- subset(longevity, select = -mass_g)
```

Do the pair plot again:

```{r}
pair_plot(longevity)
```

Split the data into a training and test set:

```{r}
# split data into training and test set
prop_train <- 0.75
s <- sample(
  c(TRUE, FALSE),
  size = n,
  replace = TRUE,
  prob = c(prop_train, 1 - prop_train)
)

train <- longevity[s, ]
test <- longevity[!s, ]

method_eff <- data.frame(
  method = character(),
  mse_tr = double(),
  mse_ts = double()
)
```

This data set contains data on the maximum lifespan of various species of
mammals and birds along with other predictors, such as mass, volancy (ability
to fly), foraging environment and so on. We will try to model the maximum
lifespan of these animals as a function of the aforementioned predictors
in the data set, excluding species and order since they are more granular
nomenclature which is unlikely to be give useful results (too many levels).

For fun, let's start with a regression tree to get a (possible not-so-accurate)
feel for the data. We will keep a track of the training and test set MSE for
the different methods employed here to see which ones are most effective in
describing the relationship in the data.

```{r}
library(tree)
longevity_tree <- tree(maximum_lifespan_yr ~ ., data = train)

tree_pred_tr <- predict(longevity_tree, train)
tree_pred_ts <- predict(longevity_tree, test)

tree_mse_tr <- mean((tree_pred_tr - train$maximum_lifespan_yr)^2)
tree_mse_ts <- mean((tree_pred_ts - test$maximum_lifespan_yr)^2)

method_eff[1, ] <- c("Regression Tree", tree_mse_tr, tree_mse_ts)
```

## Boosting

Perform boosting and optimise the shrinkage hyperparameter, \(\lambda\).
To prevent trees from becoming too convoluted I set `interaction.depth`
to 5.

```{r cache = TRUE}
library(gbm3)
library(pbapply)

n_trees <- 1000
lambda <- 10^seq(-10, -0.2, by = 0.1)

res_boost <- pblapply(lambda, function(l) {
  longevity_boost_l <- gbm(
    maximum_lifespan_yr ~ .,
    data = train,
    distribution = "gaussian",
    n.trees = n_trees,
    interaction.depth = 5,
    shrinkage = l
  )

  boost_l_pred_tr <- predict(longevity_boost_l, train, n.trees = n_trees)
  boost_l_pred_ts <- predict(longevity_boost_l, test, n.trees = n_trees)

  boost_l_mse_tr <- mean((boost_l_pred_tr - train$maximum_lifespan_yr)^2)
  boost_l_mse_ts <- mean((boost_l_pred_ts - test$maximum_lifespan_yr)^2)

  return(c(l, boost_l_mse_tr, boost_l_mse_ts))
})
```

Analysis

```{r}
boost_mse <- data.frame(do.call(rbind, res_boost))
colnames(boost_mse) <- c("lambda", "train_mse", "test_mse")

plot(
  boost_mse$lambda,
  boost_mse$train_mse,
  type = "l",
  lwd = 1.5,
  ylim = range(boost_mse$train_mse, boost_mse$test_mse),
  xlab = expression(lambda),
  ylab = "MSE"
)

lines(boost_mse$lambda, boost_mse$test_mse, lwd = 1.5, lty = 2)

legend(
  "topright",
  bty = "n",
  legend = c("Training set MSE", "Test set MSE"),
  lty = 1:2,
  lwd = 1.5
)
```

The training set error tends to zero as \(\lambda\) increases, but the
test error has a minima close to zero, due to overfitting.

```{r}
ind <- which.min(boost_mse$test_mse)
boost_lambda_opt <- boost_mse$lambda[ind]
boost_opt_mse_tr <- boost_mse$train_mse[ind]
boost_opt_mse_ts <- boost_mse$test_mse[ind]

method_eff[2, ] <- c("Boosting", boost_opt_mse_tr, boost_opt_mse_ts)
```

The optimal value for the shrinkage parameter is 
\(\lambda = `r boost_lambda_opt`\). This achieves a training set MSE of 
`r boost_opt_mse_tr` and a test set MSE of `r boost_opt_mse_ts`.

## Bagging

It's in the bag!

```{r}
library(randomForest)

longevity_bag <- randomForest(
  maximum_lifespan_yr ~ .,
  data = train,
  mtry = 6,
  importance = TRUE
)

longevity_bag
```

Compute the MSE of the training and test set (we already have the traning MSE
in the output, but I want the exact quantity).

```{r}
bag_pred_ts <- predict(longevity_bag, test)

bag_mse_tr <- mean((longevity_bag$predicted - train$maximum_lifespan_yr)^2)
bag_mse_ts <- mean((bag_pred_ts - test$maximum_lifespan_yr)^2)

method_eff[3, ] <- c("Bagging", bag_mse_tr, bag_mse_ts)
```

Importance:

```{r}
importance(longevity_bag)

varImpPlot(longevity_bag)
```

The most important variables are `log_mass_g`, `foraging_environment` and
`daily_activity`.

## Random Forest

We use random forests with hyperparameter optimisation for \(m\), the number
of variables considered at each split.

```{r}
rf_out <- sapply(1:6, function(i) {

  longevity_rf_i <- randomForest(
    maximum_lifespan_yr ~ .,
    data = train,
    mtry = i,
    importance = TRUE
  )

  rf_i_pred <- predict(longevity_rf_i, test)
  rf_i_mse <- mean((rf_i_pred - test$maximum_lifespan_yr)^2)

})

plot(1:6, rf_out, type = "b", xlab = "m", ylab = "Test set MSE")
```

Optimal value of \(m = 3\).

```{r}
longevity_rf <- randomForest(
  maximum_lifespan_yr ~ .,
  data = train,
  mtry = 3,
  importance = TRUE
)

rf_pred_ts <- predict(longevity_rf, test)

rf_mse_tr <- mean((longevity_rf$predicted - train$maximum_lifespan_yr)^2)
rf_mse_ts <- mean((rf_pred_ts - test$maximum_lifespan_yr)^2)

method_eff[3, ] <- c("Random Forest", rf_mse_tr, rf_mse_ts)
```

## BART SIMPSON

```{r}
library(BART)

x_train <- train[, -2]
y_train <- train[, 2]

x_test <- test[, -2]
y_test <- test[, 2]

longevity_bart <- gbart(x_train, y_train, x.test = x_test)
```

Calculate error:

```{r}
bart_pred_tr <- longevity_bart$yhat.train.mean
bart_pred_ts <- longevity_bart$yhat.test.mean

bart_mse_tr <- mean((bart_pred_tr - y_train)^2)
bart_mse_ts <- mean((bart_pred_ts - y_test)^2)

method_eff[4, ] <- c("BART", bart_mse_tr, bart_mse_ts)
```

Variable importances:

```{r}
ord <- order(longevity_bart$varcount.mean, decreasing = TRUE)
longevity_bart$varcount.mean[ord]
```

The most important variables according to BART are `log_mass_g`,
`foraging_environment4` (terrestrial) and `daily_activity3` (diurnal).

## Linear models

Basic linear model:

```{r}
longevity_lm <- lm(maximum_lifespan_yr ~ ., data = train)

lm_pred_tr <- longevity_lm$fitted.values
lm_pred_ts <- predict(longevity_lm, test)

lm_mse_tr <- mean((lm_pred_tr - train$maximum_lifespan_yr)^2)
lm_mse_ts <- mean((lm_pred_ts - test$maximum_lifespan_yr)^2)

method_eff[5, ] <- c("Linear model", lm_mse_tr, lm_mse_ts)
```

## Conclusions

```{r echo = FALSE}
library(knitr)

method_eff$mse_tr <- as.double(method_eff$mse_tr)
method_eff$mse_ts <- as.double(method_eff$mse_ts)
ord <- order(method_eff$mse_ts)

method_eff <- method_eff[ord, ]
rownames(method_eff) <- 1:5

kable(method_eff)
```

Boosting is performing best out of all the methods explored on this data set.
