---
title: Exercise 8.10
date: March 2024
output: html_document
author: Kári Hlynsson
---

We now use boosting to predict `Salary` in the `Hitters` data set.

## Part (a)
Remove the observations for whom the salary information is unknown, and then
log-transform the salaries.

```{r}
library(ISLR2)
data(Hitters)

hitters <- Hitters[!is.na(Hitters$Salary), ]
hitters$SalaryLog <- log(hitters$Salary)

n <- nrow(hitters)
p <- ncol(hitters)
```

## Part (b)
Create a training set consisting of the first 200 observations, and a test set
consisting of the remaining observations.

```{r}
train <- hitters[1:200, ]
test <- hitters[201:263, ]
```

## Part (c)
Perform boosting on the training set with 1,000 trees for a range of values of
the shrinkage parameter \(\lambda\). Produce a plot with different shrinkage
values on the \(x\)-axis and the corresponding training set MSE on the
\(y\)-axis.

```{r}
library(gbm3)

set.seed(1)

lambda <- 10^seq(-10, -0.2, by = 0.1)

out <- lapply(lambda, function(l) {
  hitters_boost_l <- gbm(
    SalaryLog ~ . - Salary,
    data = train,
    distribution = "gaussian",
    n.trees = 1000,
    shrinkage = l,
    verbose = FALSE
  )

  boost_l_pred_tr <- predict(hitters_boost_l, train, n.trees = 1000)
  boost_l_pred_ts <- predict(hitters_boost_l, test, n.trees = 1000)
  boost_l_mse_tr <- mean((boost_l_pred_tr - train$SalaryLog)^2)
  boost_l_mse_ts <- mean((boost_l_pred_ts - test$SalaryLog)^2)

  return(c(boost_l_mse_tr, boost_l_mse_ts))
})

lambda_mse <- do.call(rbind, out)

train_mse <- lambda_mse[, 1]
test_mse <- lambda_mse[, 2]
```

See part (d) for plot of training MSE.

## Part (d)
Produce a plot with different shrinkage values on the \(x\)-axis and the
corresponding test set MSE on the \(y\)-axis.

```{r}
dot_scale <- 0.75

plot(
  lambda,
  train_mse,
  type = "b",
  col = "blue",
  cex = dot_scale,
  xlab = expression(lambda),
  ylab = "Training set MSE",
  lwd = 1.5
)

lines(
  lambda,
  test_mse,
  type = "b",
  col = "red",
  cex = dot_scale,
  lwd = 1.5
)

legend(
  "topright",
  bty = "n",
  legend = c("Train MSE", "Test MSE"),
  col = c("blue", "red"),
  cex = dot_scale,
  lwd = 1.5
)
```

The training set MSE is decreasing with increasing values of \(\lambda\)
whereas the test set MSE decreases initially but starts increasing for
\(\lambda > 0.1\), likely due to overfitting. The optimal value for the
test set MSE is

```{r}
lambda_opt <- lambda[which.min(test_mse)]
```

or `r lambda_opt`. Fit this model for later use:

```{r}
hitters_boost <- gbm(
  SalaryLog ~ . - Salary,
  data = train,
  distribution = "gaussian",
  n.trees = 1000,
  shrinkage = lambda_opt
)

boost_mse_tr <- min(train_mse)
boost_mse_ts <- min(test_mse)
```

## Part (e)
Compare the test MSE of boosting to the test MSE that results from applying two
of the regression approaches seen in Chapters 3 and 6.

Standard linear model:

```{r}
hitters_lm <- lm(SalaryLog ~ . - Salary, data = train)

lm_pred_tr <- predict(hitters_lm, train)
lm_mse_tr <- mean((lm_pred_tr - train$SalaryLog)^2)

lm_pred_ts <- predict(hitters_lm, test)
lm_mse_ts <- mean((lm_pred_ts - test$SalaryLog)^2)
```

Training set MSE for standard linear model is `r lm_mse_tr` and the test set
MSE is `r lm_mse_ts`, both higher than the counterparts provided by boosting
regression trees.

LASSO model with CV tuning for shrinkage parameter:

```{r}
library(glmnet)

l_grid <- 10^seq(10, -2, length = 100)

x_train <- model.matrix(SalaryLog ~ ., data = train[, -19])
y_train <- train$SalaryLog

x_test <- model.matrix(SalaryLog ~ ., data = test[, -19])
y_test <- test$SalaryLog

hitters_lasso <- glmnet(x_train, y_train, alpha = 0, lambda = l_grid)
plot(hitters_lasso)
```

Find the optimal shrinkage parameter value:

```{r}
hitters_lasso_cv <- cv.glmnet(x_train, y_train, alpha = 1)
plot(hitters_lasso_cv)
```

Best \(\lambda\):

```{r}
lambda_opt <- hitters_lasso_cv$lambda.min

lasso_cv_pred_tr <- predict(hitters_lasso_cv, s = lambda_opt, x_train)
lasso_cv_mse_tr <- mean((lasso_cv_pred_tr - y_train)^2)

lasso_cv_pred_ts <- predict(hitters_lasso_cv, s = lambda_opt, x_test)
lasso_cv_mse_ts <- mean((lasso_cv_pred_ts - y_test)^2)
```

The training set MSE for an optimally-tuned LASSO regression model is
`r lasso_cv_mse_tr` and the test set MSE is `r lasso_cv_mse_ts`, again
both higher than that of the boosting approach.

This might suggest that the true relationship does not conform to a linear
model and is more suitably approximated by non-linear models.

## Part (f)
Which variables appear to be the most important predictors in the boosted
model?

```{r}
summary(hitters_boost)
```

The top 3 most important variables are `CAtBat`, `CRuns` and `CRBI`.

## Part (g)
Now apply bagging to the training set. What is the test MSE for this approach?

```{r}
library(randomForest)

hitters_bagging <- randomForest(
  SalaryLog ~ . - Salary,
  data = train,
  mtry = 19
)

bagging_pred <- predict(hitters_bagging, test)
bagging_mse_ts <- mean((bagging_pred - test$SalaryLog)^2)
```

The MSE for bagging is `r bagging_mse_ts`, which is lower than the boosting
test MSE.
