---
title: Exercise 11.11
date: April 2024
output: html_document
author: Kári Hlynsson
---

This example make use of the data in Table 11.4.

## Part (a)
Create two groups of observations. In Group 1, \(X < 2\), whereas in Group 2,
\(X \geq 2\). Plot the Kaplan-Meier survival curves corresponding to the two
groups. Be sure to label the curves so that it is clear which curve corresponds
to which group. By eye, does there appear to be a difference between the two
groups' survival curves?

```{r message = FALSE, warning = FALSE}
library(survival)

dat <- data.frame(
  Y = c(26.5, 37.2, 57.3, 90.8, 20.2, 89.8),
  d = c(1, 1, 1, 0, 0, 0),
  X = c(0.1, 11, -0.3, 2.8, 1.8, 0.4)
)

dat$G <- as.factor(ifelse(dat$X < 2, "1", "2"))

attach(dat)

surv_fit <- survfit(Surv(Y, d) ~ G)
plot(surv_fit, lwd = 2.25, col = 2:3, xlab = "Y", ylab = "Pr(d = 0 | Y > y)")
legend("topright", bty = "n", legend = levels(G), lwd = 2.25, col = 2:3)
```

No discernible difference at first glance, or not one I would be comfortable
standing by.

## Part (b)
Fit Cox's proportional hazards model, using the group indicator as a covariate.
What is the estimated coefficient? Write a sentence providing the
interpretation of this coefficient, in terms of the hazard or the
instantaneous probability of the event. Is there evidence that the true
coefficient value is non-zero?

```{r}
cox_fit <- coxph(Surv(Y, d) ~ G)
summary(cox_fit)
```

Estimated coefficient shown in output above. The interpretation is that for an
observation identical to another in every regard other than being G2 whereas
its counterpart is G1, this would correspond to them being 
\(exp(-0.3401) = 0.7117\) less likely to experience the event. However, the
Wald test is not significant \(p \approx 0.783\) and thus we can not sustain
this claimin actuality.

## Part (c)
Recall from Section 11.5.2 that in the case of a single binary covariate, the
log-rank test statistic should be identical to the score statistic for the Cox
model. Conduct a log-rank test to deterine whether there is a difference
between the survival curves for the two groups. How does the \(p\)-value for
the log-rank test statistic compare to the previous \(p\)-value for the score
statistic for the Cox model from (b)?

```{r}
summary(cox_fit)$sctest
survdiff(Surv(Y, d) ~ G)$pvalue
```

\(p\)-values insignificant in both tests -- no evidence for claiming that there
is a difference in outcome between groups G1 and G2.
