---
title: Exercise 12.13
date: April 2024
output: html_document 
author: Kári Hlynsson
---

On the book website, `www.statlearning.com`, there is a gene expression data
set (`Ch12Ex13.csv`) that consists of 40 tissue samples with measurements on
1,000 genes. The first 20 samples are from healthy patients, while the second
20 are from a diseased group.

## Part (a)
Load in the data using `read.csv()`. You will need to select `header = F`.

```{r}
tmp <- tempfile()
download.file(
  url = "https://www.statlearning.com/s/Ch12Ex13.csv",
  dest = tmp
)
x <- t(read.csv(tmp, header = FALSE))
rm(tmp)

# 1 = healthy, 2 = diseased
status <- c(rep(1, 20), rep(2, 20))
```

Data cleaning:

```{r}
rownames(x) <- paste0("P", 1:40)
colnames(x) <- paste0("G", 1:1000)
```

Let's compute some summary statistics to infer whether scaling is necessary:

```{r}
apply(x, 2, mean) |> range()
```

Variances:

```{r}
apply(x, 2, var) |> range()
```

Close enough -- I don't think scaling is necessary.

## Part (b)
Apply hierarchical clustering to the samples using correlation-based distance,
and plot the dendrogram. Do the genes separate the samples in to the two
groups? Do your results depend on the type of linkage used?

```{r}
dd <- as.dist(1 - cor(t(x)))

dclust <- function(link) {
  xclust <- hclust(dd, method = link)
  plot(xclust)

  pred <- cutree(xclust, 2)

  return(table(pred, status))
}

cl_complete <- dclust("complete")
```

Split dendrogram based on the two groups:

```{r}
cl_complete
```

The hierarchical clustering correctly identifies all diseased patients, but
confuses almost half the healthy patients in case of complete linkage. Let's
try single linkage.

```{r}
cl_single <- dclust("single")
cl_single
```

Single linkage is better in terms of identifying healthy patients, but abysmal
when it comes to identifying diseased patients -- in fact, it has a 100% error
rate when it comes to identifying them.

Average linkage:

```{r}
cl_average <- dclust("average")
cl_average
```

Similar but slightly worse performance to the complete linkage clustering.

Lastly, we have centroid linkage:

```{r}
cl_centroid <- dclust("centroid")
cl_centroid
```

Terrible. Complete linkage turned out to be the best method.

## Part (c)
Your collaborator wants to know which genes differ the most across the two
groups. Suggest a way to answer this question, and apply it here.

As seen in (b), using hierarchical clustering with completel linkage yielded
the best results where 100% of diseased patients were correctly identified.
However, healthy patients were often misclassified which is not so good.

We might want to test whether we can use PCA to identify genes central to the
expression of disease status. I.e., if the principal components divide the
two groups visually, we might be in luck. Let's try it out:

```{r}
pr <- prcomp(x, scale = TRUE)

biplot(pr, scale = 0)
```

Although the principal components are fairly indecisive, there appears to be
some division in the data when projected onto the principal component score
values -- might these be our groups?

```{r}
plot(pr$x[, 1], pr$x[, 2], pch = paste(status), col = status + 2, cex = 1.5,
     xlab = "PC1", ylab = "PC2")
```

Jackpot! We can classify the patients in terms of gene expression using the
first principal component:

```{r}
pred_pca <- ifelse(pr$x[, 1] > 0, 2, 1)

table(pred_pca, status)
```

Yippee!
