---
title: Exercise 12.9
date: April 2024
output: html_document
author: Kári Hlynsson
---

Consider the `USArrests` data. We will now perform hierarchical clustering on
the states.

## Part (a)
Using hierarchical clustering with complete linkage and Euclidean distance,
cluster the states.

```{r}
library(ISLR2)
data(USArrests)

hclust_unsc <- hclust(dist(USArrests), method = "complete")
```

Let's take a look.

```{r}
plot(hclust_unsc, main = "Hierarchical clustering (unscaled)",
     xlab = "State")
abline(h = 150, lty = 2, col = "red")
```

## Part (b)
Cut the dendogram at a height that results in three distinct clusters. Which
states belong to which clusters?

```{r}
clusters_unsc <- cutree(hclust_unsc, 3)

clusters_unsc
table(clusters_unsc)
```

## Part (c)
Hierarchically cluster the states using complete linkage and Euclidean
distance, *after scaling the variables to have standard deviation one*.

```{r}
xsc <- scale(USArrests)

hclust_sc <- hclust(dist(xsc), method = "complete")

plot(hclust_sc, main = "Hierarchical clustering (scaled)")
abline(h = 4, lty = 2, col = "red")
```

Let's take a look at how the cluster groups look like:

```{r}
clusters_sc <- cutree(hclust_sc, 4)

clusters_sc
table(clusters_sc)
```

## Part (d)
What effect does scaling the variales have on the hierarchical clustering
obtained? In your opinion, should the variables be scaled before the
inter-observation dissimilarities are computed? Provide a justification for
your answer.

We see that the dendrogram for the scaled variables is more balanced in the
sense that we have four large subtrees, instead of the three produced by the
unscaled response. It remains to be answered whether the data would suggest
a clustering into 3 or 4 states -- perhaps they are related by geography,
political tendencies etc. I would likely prefer to scale the data, but I would
need to look at how the groups change and whether there is any structure in
either of the two clusterings produced.

It would be interesting to plot the clusters on a map of the USA to see whether
we see any discernible patterns in the clustering.
