---
title: Exercise 12.8
date: April 2024
output: html_document
author: Kári Hlynsson
---

In Section 12.2.3, a formula for calculating PVE was given in Equation 12.10.
We also saw that the `PVE` can be obtained using the `sdev` output of the
`prcomp()` function.

On the `USArrests` data, calculate PVE in two ways:

## Part (a)
Using the `sdev` output of the `prcomp()` function, as was done in Section
12.2.3.

```{r}
library(ISLR2)
data(USArrests)

n <- nrow(USArrests)
p <- ncol(USArrests)

pr <- prcomp(USArrests, scale = TRUE)
biplot(pr, scale = 0, cex = 0.75)
```

Calculate the PVE:

```{r}
pr_var <- pr$sdev^2
pve <- pr_var / sum(pr_var)
names(pve) <- paste0("PC", 1:4)

pve
```

Two neat plots:

```{r}
par(mfrow = c(1, 2))

plot(
  1:4, pve, xlab = "PC", ylab = "PVE",
  ylim = c(0, 1), type = "b", xaxt = "n"
)

axis(side = 1, at = 1:4)

plot(
  1:4, cumsum(pve), xlab = "PC", ylab = "Cumulative PVE",
  ylim = c(0, 1), type = "b", xaxt = "n"
)

axis(side = 1, at = 1:4)
```

## Part (b)
By applying Equation 12.10 directly. That is, use the `prcomp()` function
to compute the principal component loadings. Then, use those loadings in
Equation 12.10 to obtain the PVE.

```{r}
x <- as.matrix(scale(USArrests))

pve2 <- sapply(1:4, function(i) {
  pc <- pr$rotation[, i]
  denom <- sum(sapply(1:n, function(i) (x[i, ] %*% pc)^2))
  numer <- sum(x^2)
  return(denom / numer)
})

names(pve2) <- paste0("PC", 1:4)
```

The results are virtually identical.

```{r}
pve
pve2
```


