---
title: Exercise 13.7
date: April 2024
output: html_document
author: Kári Hlynsson
---

This problem makes use of the `Carseats` data set in the `ISLR2` package.

## Part (a)
For each quantitative variable in the data set besides `Sales`, fit a linear
model to predict `Sales` using that quantitative variable. Report the
\(p\)-values associated with the coefficients for the variables. That is, for
each model of the form \(Y = \beta_0 + \beta_1 X + \epsilon\), report the
\(p\)-value associated with the coefficient \(\beta_1\). Here, \(Y\) represents
`Sales` and \(X\) represents one of the other quantitative variables.

```{r}
library(ISLR2)
data(Carseats)
```

Fit the models and obtain the \(p\)-values:

```{r}
vars <- c("CompPrice", "Income", "Advertising", "Population", "Price",
          "Age", "Education")

pvals <- sapply(vars, function(var) {
  fit <- lm(as.formula(paste0("Sales ~ ", var)), data = Carseats)
  pval <- summary(fit)$coefficients[2, 4]
  return(pval)
})
```

## Part (b)
Suppose we control the Type I error at level \(\alpha = 0.05\) for the
\(p\)-values obtained in (a). Which null hypotheses do we reject?

```{r}
names(which(pvals < 0.05))
```

`Income`, `Advertising`, `Price` and `Age`.

## Part (c)
Now suppose we control the FWER at level 0.05 for the \(p\)-values. Which null
hypotheses do we reject?

```{r}
# Bonferroni
names(which(p.adjust(pvals, method = "bonferroni") < 0.05))

# Holm
names(which(p.adjust(pvals, method = "holm") < 0.05))
```

## Part (d)
Finally, suppose we control the FDR at level 0.2 for the \(p\)-values. Which
null hypotheses do we reject?

```{r}
names(which(p.adjust(pvals, method = "fdr") < 0.2))
```

Always the same variables, namely `Income`, `Advertising`, `Price` and `Age`.
