---
title: Exercise 13.7
date: April 2024
output: html_document
author: Kári Hlynsson
---

This problem makes use of the `Carseats` data set in the `ISLR2` package.

## Part (a)
For each quantitative variable in the data set besides `Sales`, fit a linear
model to predict `Sales` using that quantitative variable. Report the
\(p\)-values associated with the coefficients for the variables. That is, for
each model of the form \(Y = \beta_0 + \beta_1 X + \epsilon\), report the
\(p\)-value associated with the coefficient \(\beta_1\). Here, \(Y\) represents
`Sales` and \(X\) represents one of the other quantitative variables.

### Solution.
We start by importing the `Carseats` data set and selecting the quantitative
predictors which are stored in the variable `vars`:

```{r}
library(ISLR2)
data(Carseats)

vars <- names(
  which(sapply(Carseats, is.numeric) & names(Carseats) != "Sales")
) |> sort()
```

Now we fit a series of simple linear models 
\(\mathcal M_1, \ldots, \mathcal M_p\) of `Sales` (\(Y\)) against each of the
quantitative covariates \(X_1, \ldots, X_p\) contained in `vars`. In other
words, we fit the series of models
\begin{align*}
  \mathcal M_1: Y &= \beta_{01} + \beta_{11} X_1 + \epsilon \\
  &\vdots \\
  \mathcal M_p: Y &= \beta_{0p} + \beta_{1p} X_p + \epsilon,
\end{align*}
and investigate the value of the Wald test statistic \(p_j\)
associated with the effect size \(\beta_{1j}\) in model \(\mathcal M_j\),
each of which corresponds to a hypotheses test of the form
\[
  H_0: \beta_{1j} = 0, \quad H_1: \beta_{1j} \neq 0.
\]
The later parts of this exercise will revolve around different methods for
controlling Type I error in light of multiple testing. Recover the \(p\)-values:

```{r}
pvals <- sapply(vars, function(var) {
  fit <- lm(as.formula(paste0("Sales ~ ", var)), data = Carseats)
  pval <- summary(fit)$coefficients[2, 4]
  return(pval)
})

pvals
```

Notice that the \(p\)-values for predictors `Advertising`, `Age`, `Income` and
`Price` are markedly small compared to the other values. This is reinforced by
the visualisation of the various predictors against `Sales`. Indeed, in the
case of the four variables, there is seemingly some discernible linearity in 
the data.

```{r}
par(mfrow = c(2, 4))
for (var in vars) {
  plot(Carseats[[var]], Carseats[["Sales"]], xlab = paste(var), ylab = "Sales")
  abline(lm(Carseats[["Sales"]] ~ Carseats[[var]]),
    lwd = 2, col = "red"
  )
  abline(h = mean(Carseats[["Sales"]]), lty = 2, col = "red")
}
```

## Part (b)
Suppose we control the Type I error at level \(\alpha = 0.05\) for the
\(p\)-values obtained in (a). Which null hypotheses do we reject?

### Solution.
This is the simple case where in a manner analogous to single hypothesis
testing we simply determine whether \(p_j < \alpha\) sequentially irrespective
of the number of tests performed. If \(p_j < \alpha\) we reject the null
hypothesis \(H_0\) in model \(\mathcal M_j\) in favour of the claim that
\(\beta_{1j} \neq 0\), i.e. that the effect size of the \(j\)th covariate
in a simple linear model is distinct from zero.

Doing this in `R` amounts to simply filtering the `pvals` vector for values
below \(\alpha = 0.05\):

```{r}
names(which(pvals < 0.05))
```

The significant tests are for the predictors `Advertising`, `Age`, `Income` 
and `Price`. (Compare with output in part (a).)

## Part (c)
Now suppose we control the FWER at level 0.05 for the \(p\)-values. Which null
hypotheses do we reject?

### Solution.
We will employ two methods that are popular for controlling the FWER, namely
the Bonferroni correction and the Holm Step-Down method, both of which are
described in detail in Chapter 13. Start by making the adjustments to `pvals`:

```{r}
p_bonf <- p.adjust(pvals, method = "bonferroni")
p_bonf

p_holm <- p.adjust(pvals, method = "holm")
p_holm
```

Indeed, we see that the \(p\)-values for the four significant covariates from
(b) remain significant.

```{r}
# Bonferroni
names(which(p_bonf < 0.05))

# Holm
names(which(p_holm < 0.05))
```

## Part (d)
Finally, suppose we control the FDR at level 0.2 for the \(p\)-values. Which
null hypotheses do we reject?

### Solution.
Lastly we employ methods to control the false discovery rate, or FDR. One
way with which we can do this is using the Benjamini-Hochberg procedure,
which is designated as `bh` or `fdr` in the `method` variable in the 
`p.adjust()` function.

```{r}
# Benjamini-Hochberg
p_bh <- p.adjust(pvals, method = "fdr")
p_bh
```

Let's see which of the predictors are significant using the Benjamini-Hochberg
adjusted \(p\)-values:

```{r}
names(which(p_bh < 0.2))
```

The results are not very novel. The same four predictors, namely `Advertising`,
`Age`, `Income` and `Price` are claimed to have significantly non-zero effect
sizes.

