---
title: "Exercise 4.14"
author: "Kári Hlynsson"
date: "2024-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MASS)
library(e1071)
library(class)
library(tidyverse)
library(caret)
library(ISLR2)
```


In this problem, you will develop a model to predict whether a given car gets high or low gas mileage based on the `Auto` data set.

## Part (a)
Create a binary variable, `mpg01`, that contains a 1 if `mpg` contains a value above its median, and a 0 if `mpg` contains a value below its median. You can compute the median using the `median()` function. Note you may find it helpful to use the `data.frame()` function to create a single data set containing both `mpg01` and the other `Auto` variables.

```{r}
data(Auto)

Auto <- Auto |> mutate(mpg01 = factor(ifelse(mpg > median(mpg), 1, 0)))
```

## Part (b)
Explore the data graphically in order to investigate the association between `mpg01` and the other features. Which of the other features seem most likely to be useful in predicting `mpg01`? Scatterplots and boxplots may be useful tools to answer this question. Describe your findings.

```{r}
GGally::ggpairs(Auto |> select(-c(name, mpg)), progress = FALSE) 

pairs(Auto |> select(-name, mpg), col = Auto$mpg01)
```

I chose not to include `mpg` in the pair plot as it is directly related to `mpg01`. Judging from the plot, `acceleration`, `weight`, `horsepower` and `displacement`
all seem to be correlated with `mpg01`. We may need to treat interaction effects between these variables though.

## Part (c)
Split the data into a training set and a test set.

```{r}
set.seed(1)

mask <- sample(1:nrow(Auto), 0.75 * nrow(Auto))

train <- Auto[mask,]
test  <- Auto[setdiff(1:nrow(Auto), mask),]
```


## Part (d)
> Perform LDA on the training data in order to predict `mpg01` using the variables that seemed most associated with `mpg01` in (b). What is the test error of the model obtained?

```{r}
lda.fit <- lda(mpg01 ~ displacement + horsepower + weight + acceleration, data = train)

lda.pred <- predict(lda.fit, test)
confusionMatrix(lda.pred$class, test$mpg01)
```

The test set error is $13 / 98 \approx 0.1327$, the complement of the accuracy as shown in the output of `confusionMatrix` above. The LDA classifier has
very good specificity, and quite good sensitivity. Let's try adding interaction effects between some of the variables.

```{r}
lda.fit.2  <- lda(mpg01 ~ acceleration + displacement + horsepower:weight, data = train)
lda.pred.2 <- predict(lda.fit.2, test)

confusionMatrix(lda.pred.2$class, test$mpg01)
```

Adding an interaction effect between `weight` and `horsepower` increases the classifier accuracy.

## Part (e)
> Perform QDA on the training data in order to predict `mpg01` using the variables that seemed most associated with `mpg01` in (b). What is the test error of the model obtained?

```{r}
qda.fit <- qda(mpg01 ~ acceleration + displacement + horsepower:weight, data = train)
qda.pred <- predict(qda.fit, test)

confusionMatrix(qda.pred$class, test$mpg01)
```

The model performs better than its LDA counterpart. Test set error is $\approx 0.0816$.

## Part (f)
Perform logistic regression on the training data in order to predict `mpg01` using the variables that seemed most associated with `mpg01` in (b). What is the test error of the model obtained?

```{r}
glm.fit   <- glm(mpg01 ~ acceleration + displacement + horsepower:weight, family = "binomial", data = train)

glm.probs <- predict(glm.fit, test, type = "response")
glm.pred  <- rep(0, nrow(test))
glm.pred[glm.probs > 0.5] <- 1

glm.pred <- as.factor(glm.pred)

confusionMatrix(glm.pred, test$mpg01)
```

Logistic regression is not as good as the previous two methods we used.

## Part (g)
> Perform naive Bayes on the training data in order to predict `mpg01` using the variables that seemed most associated with `mpg01` in (b). What is the test error of the model obtained?

Naive Bayes is unable to accomodate interaction terms, so we will use the covariates without any such adjustments.

```{r}
nb.fit <- naiveBayes(mpg01 ~ acceleration + displacement + horsepower + weight, data = train)
nb.pred <- predict(nb.fit, test)

confusionMatrix(nb.pred, test$mpg01)
```

Does not perform as well.

## Part (h)
> Perform KNN on the training data, with several values of K, in order to predict `mpg01`. Use only the variables that seemed most associated with `mpg01` in (b). What test errors do you obtain? Which value of K seems to perform the best on this data set?

```{r}
train.X <- train |> select(acceleration, displacement, horsepower, weight)
test.X  <- test |> select(acceleration, displacement, horsepower, weight)

train.Y <- train |> pull(mpg01)

knn.pred  <- knn(train.X, test.X, train.Y, k = 3)
```

Construct the confusion matrix:

```{r}
confusionMatrix(knn.pred, test$mpg01)
```

Now let's tune the hyperparameters:

```{r}
k.accuracy <- sapply(1:30, function(k) {
  l <- c()
  for (i in 1:1000) {
    knn.pred <- knn(train.X, test.X, train.Y, k)
    accuracy <- confusionMatrix(knn.pred, test$mpg01)$overall[["Accuracy"]]
    l <- append(l, accuracy)
  }
  mean(l)
})
```

Find the optimal value of \(K\):

```{r}
which.max(k.accuracy)
```

It's 4.