---
title: "Exercise 6.9"
author: "Kári Hlynsson"
date: "2024-02-12"
output: html_document
---

In this exercise, we will predict the number of applications received using the other variables in the `College` data set.

## Part (a)
Split the data into a training set and a test set.

```{r}
library(ISLR2)
data(College)

set.seed(42)
n <- nrow(College)
s <- sample(c(TRUE, FALSE), n, replace = TRUE, prob = c(0.75, 0.25))

# for lm
train <- College[s,]
test  <- College[!s,]

# for glmnet
model.mat <- model.matrix(Apps ~ ., data = College)
train.X   <- model.mat[s,  -c(1, 3)]
train.Y   <- model.mat[s,         3]
test.X    <- model.mat[!s, -c(1, 3)]
test.Y    <- model.mat[!s,        3]
```


## Part (b)
Fit a linear model using least squares on the training set, and report the test error obtained.

```{r}
lm.fit  <- lm(Apps ~ ., data = train)
lm.pred <- predict(lm.fit, test)

lm.err  <- mean((test$Apps - lm.pred)^2)
lm.err # MSE
```

## Part (c)
Fit a ridge regression model on the training set, with $\lambda$ chosen by cross-validation. Report the test error obtained.

```{r}
library(glmnet)

grid      <- 10^seq(10, -2, length = 100)
ridge.mod <- glmnet(train.X, train.Y, alpha = 0, lambda = grid)
plot(ridge.mod)
```

Weird. Find optimal $\lambda$ for ridge regression:

```{r}
cv.out     <- cv.glmnet(train.X, train.Y, alpha = 0)
lambda.opt <- cv.out$lambda.min
```

Now find the model coefficients:

```{r}
ridge.coefs <-predict(ridge.mod, s = lambda.opt, type = "coefficients")
barplot(mod.coefs[,1])
```

Measure model performance on test data set:

```{r}
ridge.pred <- predict(ridge.mod, s = lambda.opt, newx = test.X)
mean((test.Y - ridge.pred)^2)
```

Ridge model outperforms least squares.

## Part (d)
Fit a lasso model on the training set, with $\lambda$ chosen by cross-validation. Report the test error obtained, along with the number of non-zero coefficient estimates.

```{r warning=FALSE}
lasso.mod <- glmnet(train.X, train.Y, alpha = 1, lambda = grid)
plot(lasso.mod)
```

Find optimal value for $\lambda$:

```{r}
cv.out     <- cv.glmnet(train.X, train.Y, alpha = 1)
lambda.opt <- cv.out$lambda.min
lambda.opt
```

Find coefficients:

```{r}
predict(lasso.mod, s = lambda.opt, type = "coefficients")
```

Measure model error on test set.

```{r}
lasso.pred <- predict(lasso.mod, s = lambda.opt, newx = test.X)
mean((lasso.pred - test.Y)^2)
```

Lasso outperforms OLS again.

## Part (e)
Fit a PCR model on the training set, with $M$ chosen by cross-validation. Report the test error obtained, along with the value of $M$
selected by cross-validation.

```{r}
library(pls)

pcr.fit <- pcr(Apps ~ ., data = train, scale = TRUE, validation = "CV")
summary(pcr.fit)
```

Validation plot.

```{r}
validationplot(pcr.fit, val.type = "MSEP")
```

We get lowest training error on 17 components, i.e. the whole data set.

## Part (f)
Fit a PLS model on the training set, with $M$ chosen by cross-validation. Report the test error obtained, along with the value of $M$
selected by cross-validation.

## Part (g)
Comment on the results obtained. How accurately can we predict the number of college applications received? Is there much difference
among the test errors resulting from these five approaches?