---
title: "Exercise 3.11"
author: "Kári Hlynsson"
date: "2024-02-13"
output: html_document
editor_options: 
  markdown: 
    wrap: 100
---

We will now try to predict per capita crime rate in the `Boston` data set.

## Part (a)

Try out some of the regression methods explored in this chapter, such as best subset selection, the
lasso, ridge regression, and PCR. Present and discuss results for the approaches that you consider.

```{r}
library(MASS)
library(ISLR2)
data(Boston)

set.seed(1)

n <- nrow(Boston)
s <- sample(n, round(0.75 * n))
t <- setdiff(1:n, s)

train <- Boston[s,]
test  <- Boston[t,]

train.X <- model.matrix(crim ~ ., data = train)[,-1]
train.Y <- train[,1]

test.X <- model.matrix(crim ~ ., data = test)[,-1]
test.Y <- test[,1]
```

### Best subset selection

```{r}
library(leaps)

regfit.full <- regsubsets(crim ~ ., data = train, nvmax = 13)
reg.summary <- summary(regfit.full)
reg.summary
```

```{r}
par(mfrow = c(2, 3))
plot(1:12, reg.summary$bic, xlab = "Model size", ylab = "BIC", type = "b")
plot(1:12, reg.summary$adjr2, xlab = "Model size", ylab = expression('adj. R'^2), type = "b")
plot(1:12, reg.summary$cp, xlab = "Model size", ylab = expression(C[p]), type = "b")
plot(regfit.full, scale = "bic")
plot(regfit.full, scale = "adjr2")
plot(regfit.full, scale = "Cp")
```

```{r}
which.min(reg.summary$bic)
which.max(reg.summary$adjr2)
which.min(reg.summary$cp)
```

I will take the subset with index 8 since it does a good job of minimising the adjusted $R^2$ and
$C_p$ while also keeping the $\mathrm{BIC}$ relatively small. We could also look at how the
different subsets perform in terms of training and test set MSE.

```{r}
bss.coefs <- coef(regfit.full, id = 8)

train.mses <- sapply(1:12, function(i) {
  beta <- coef(regfit.full, id = i)
  vars  <- which(colnames(Boston) %in% names(beta))
  
  # predict Y
  X.pred    <- as.matrix(cbind(1, Boston[s,vars]))
  Y.pred    <- X.pred %*% beta
  
  # MSE
  mean((train$crim - Y.pred)^2)
})

test.mses <- sapply(1:12, function(i) {
  beta <- coef(regfit.full, id = i)
  vars  <- which(colnames(Boston) %in% names(beta))
  
  # predict Y
  X.pred    <- as.matrix(cbind(1, Boston[t,vars]))
  Y.pred    <- X.pred %*% beta
  
  # MSE
  mean((test$crim - Y.pred)^2)
})

par(mfrow = c(1, 2))
plot(train.mses, type = "b", xlab = "Model size", ylab = "Training set MSE")
plot(test.mses, type = "b", xlab = "Model size", ylab = "Test set MSE")
```

Which model performs best according to the test and training MSE?

```{r}
which.min(train.mses)
which.min(test.mses)
```

Empirical data suggests that the model of size 11 is most suitable. However I would stick with the
model of size 8 since it has lower adjusted $R^2$, $C_p$ and higher $\mathrm{BIC}$. Then the test
set MSE is

```{r}
bss.mse <- test.mses[8]
bss.mse
```

### Stepwise regression

We'll do bidirectional stepwise regression, starting from the full model:

```{r}
lm.full <- lm(crim ~ ., data = train)
lm.aic  <- stepAIC(lm.full, direction = "both", trace = FALSE)
summary(lm.aic)
```

The returned model uses 8 predictors, similar to the model we chose in best subset selection.
Measure performance:

```{r}
lm.aic.pred <- predict(lm.aic, test)
lm.aic.mse  <- mean((lm.aic.pred - test.Y)^2)
lm.aic.mse
```

Performs similarly to rest.

### Ridge regression

```{r message = FALSE}
library(glmnet)

grid <- 10^seq(-2, 10, length = 100)
ridge.fit    <- glmnet(train.X, train.Y, lambda = grid, alpha = 0, thresh = 1e-12)
cv.out.ridge <- cv.glmnet(train.X, train.Y, lambda = grid, alpha = 0, thresh = 1e-12)
plot(cv.out.ridge)
```

What is the optimal $\lambda$ selected by `cv.glmnet`?

```{r}
ridge.lambda <- cv.out.ridge$lambda.min
ridge.lambda
```

It is quite small, so it favours an OLS-esque model. Let's take a look at the model produced with
this value of $\lambda$.

```{r}
ridge.coefs <- predict(ridge.fit, s = ridge.lambda, type = "coefficients")
ridge.coefs
```

It is very similar to a OLS fit:

```{r}
summary(lm(crim ~ ., data = train))
```

Let's get a sense for model performance:

```{r}
ridge.pred <- predict(ridge.fit, s = ridge.lambda, newx = test.X)
ridge.mse  <- mean((ridge.pred - test.Y)^2)
ridge.mse
```

The performance is similar to that of best subset selection.

### Lasso regression

```{r warning = FALSE}
lasso.fit <- glmnet(train.X, train.Y, alpha = 1, lambda = grid, thresh = 1e-12)
plot(lasso.fit)
```

Optimise $\lambda$:

```{r}
cv.out.lambda <- cv.glmnet(train.X, train.Y, alpha = 1, lambda = grid, thresh = 1e-12)
lasso.lambda  <- cv.out.lambda$lambda.min
lasso.lambda
```

Again the chosen value of $\lambda$ is close to zero, so the shrinkage is very little.

```{r}
lasso.coefs <- predict(lasso.fit, s = lasso.lambda, type = "coefficients")
lasso.coefs
```

Again very similar to the OLS results.

```{r}
lasso.pred <- predict(lasso.fit, s = lasso.lambda, newx = test.X)
lasso.mse  <- mean((lasso.pred - test.Y)^2)
lasso.mse
```

### PCR

```{r}
library(pls)

pcr.fit <- pcr(crim ~ ., data = train, scale = TRUE, validation = "CV")
summary(pcr.fit)
```

PCR wants to keep all 12 components in the model.

```{r}
validationplot(pcr.fit, val.type = "MSEP")
```

```{r}
pcr.pred <- predict(pcr.fit, test)
pcr.mse  <- mean((pcr.pred - test.Y)^2)
pcr.mse
```

Performs only slightly worse than other alternatives in this instance.

### PLS

```{r}
pls.fit <- plsr(crim ~ ., data = train, scale = TRUE, validation = "CV")
summary(pls.fit)
```

From the output we see that $M = 11$ has minimal validation error, so we'll use that:

```{r}
validationplot(pls.fit, val.type = "MSEP")
```

```{r}
pls.pred <- predict(pls.fit, test)
pls.mse  <- mean((pls.pred - test.Y)^2)
pls.mse
```

## Part (b)

Propose a model (or set of models) that seem to perform well on this data set, and justify your
answer. Make sure that you are evaluating model performance using validation set error,
cross-validation, or some other reasonable alternative, as opposed to using training error.

```{r}
names <- c("Best subset", "Step AIC", "Ridge", "Lasso", "PCR", "PLS")
msev  <- c(bss.mse, lm.aic.mse, ridge.mse, lasso.mse, pcr.mse, pls.mse)
mses  <- setNames(msev, names)

barplot(mses, cex.names = 0.8)
```

Compute the $R^2$ to compare all methods:

```{r}
mtss <- mean((test.Y - mean(test.Y))^2)

rsqs <- sapply(mses, function(mse) 1 - mse / mtss)
barplot(rsqs, cex.names = 0.85)
```

Which is optimal? It's hard to tell from the visuals. :-)

```{r}
which.max(rsqs)
```

Best subset is performing best on the data in this instance.

## Part

Does your chosen model involve all of the features in the data set? Why or why not?

```{r}
bss.coefs
```

The model includes 8 predictors. First of all we chose the "optimal" model from the best subset
selection using a mixture of metrics, including BIC, adjusted $R^2$ and $C_p$ each of which have
their advantages and disadvantages. The chosen subset was picked to optimise all of these.
