---
title: "Exercise 5.9"
author: "Kári Hlynsson"
date: "2024-01-21"
output: html_document
---

We will now consider the `Boston` housing data set, from the `ISLR2` library.

## Part (a)

Based on this data set, provide an estimate for the population mean of `medv`. Call this estimate $\hat\mu$.

```{r message = FALSE}
library(ISLR2)
data(Boston)
attach(Boston)

n <- nrow(Boston)
p <- ncol(Boston)
```

```{r}
hat_mu <- mean(medv)
hat_mu
```

## Part (b)

Provide an estimate of the standard error of $\hat\mu$. Interpret this result.

*Hint: We can compute the standard error of the sample mean by dividing the sample standard deviation by the square root of the number of observations.*

By the central limit theorem, we have that the distribution of the mean of `medv` given a random sample is normal with mean $\hat\mu$ and standard deviation $\sigma/\sqrt{n}$. Thus we can approximate this as follows:

```{r}
se_hat_mu <- sd(medv) / sqrt(n)
se_hat_mu
```

## Part (c)

Now estimate the standard error of $\hat\mu$ using the bootstrap. How does this compare to your answer from (b)?

```{r}
library(boot)

boot.fn <- function(data, index) mean(data[index])
boot.res <- boot(medv, boot.fn, R = 10000)

boot.res
```

The estimate is very similar to the one from before.

## Part (d)

Based on your bootstrap estimate from (c), provide a 95% confidence interval for the mean of `medv`. Compare it to the results obtained using `t.test(Boston$medv)`.

*Hint: You can approximate a 95% confidence interval using the formula* $[\hat\mu - 1.96 \cdot \mathrm{SE}(\hat\mu), \hat\mu + 1.96 \cdot \mathrm{SE}(\hat\mu)]$.

```{r}
se_boot <- sd(boot.res$t)

ci_t <- t.test(medv)$conf.int
ci_b <- c(hat_mu - se_boot, hat_mu + se_boot)
```

## Part (e)

Based on this data set, provide an estimate, $\hat\mu_{\mathrm{med}}$, for the median value of `medv` in the population.

```{r}
hat_median <- median(medv)
hat_median
```

## Part (f)

We now would like to estimate the standard error of $\hat\mu_{\mathrm{med}}$. Unfortunately, there is no simple formula for computing the standard error of the median. Instead, estimate the standard error of the median using the bootstrap. Comment on your findings.

```{r}
boot.fn  <- function(data, index) median(data[index])
boot.res <- boot(medv, boot.fn, R = 10000)
boot.res
```

The estimated standard error is 0.3778103.

## Part (g)

Based on this data set, provide an estimate for the tenth percentile of `medv` in Boston census tracts. Call this quantity $\hat\mu_{0.1}$. (You can use the `quantile()` function.)

```{r}
quantile(medv, 0.1)
```

## Part (h)

Use the bootstrap to estimate the standard error of $\hat\mu_{0.1}$. Comment on your findings.

```{r}
boot.fn <- function(data, index) quantile(data[index], 0.1)
boot.res <- boot(medv, boot.fn, R = 10000)
boot.res
```

The standard error is 0.5075533.
