---
title: "Week 3 — Exercise 5.6"
author: "Kári Hlynsson"
output: 
  bookdown::html_document2:
    toc: true
    toc_float:
      toc_collapsed: false    
      smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

## Packages

```{r}
library(ISLR2)
library(caret)
library(tidyverse)

data(Default)

set.seed(1) # :D
```

We continue to consider the use of a logistic regression model to predict the probability of `default` using `income` and `balance` on the `Default` data set. In particular, we will now compute estimates for the standard errors of the `income` and `balance` logistic regression coefficients in two different ways: (1) using the bootstrap, and (2) using the standard formula for computing the standard errors in the `glm()` function. Do not forget to set a random seed before beginning your analysis.

## Part (a)
Using the `summary()` and `glm()` functions, determine the estimated standard errors for the coefficients associated with income and balance in a multiple logistic regression model that uses both predictors.

```{r}
init.fit <- glm(default ~ balance + income, data = Default, family = "binomial")
summary(init.fit)
```

The standard deviation for `balance` is $5.647 \cdot 10^{-3}$ is standard error for `income` is

