---
title: "Exercise 5.6"
author: "Kári Hlynsson"
date: "2024-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

We consider to consider the use of a logistic regression model to predict the probability of `default` using `income` and `balance` on the `Default` data set. In particular, we will now compute estimates for the standard errors of the `income` and `balance` logistic regression coefficients in two different ways: (1) using the bootstrap, and (2) using the standard formula for computing the standard errors in the `glm()` function. Do not forget to set a random seed before beginning your analysis.

## Part (a)

Using the `summary()` and `glm()` functions, determine the estimated standard errors for the coefficients associated with `income` and `balance` in a multiple logistic regression model that uses both predictors.

```{r message = FALSE}
library(ISLR2)

data(Default)
attach(Default)

glm.fit <- glm(default ~ income + balance, family = "binomial")
summary(glm.fit)
```

## Part (b)

Write a function, `boot.fn()`, that takes as input the `Default` data set as well as an index of observations, and that outputs the coefficient estimates for `income` and `balance` in the multiple logistic regression model.

```{r}
library(boot)

boot.fn <- function(data, index) {
  fit <- glm(default ~ income + balance, family = "binomial", subset = index)
  coef(fit)
}
```

## Part (c)

Use the `boot()` function together with your `boot.fn()` function to estimate the standard errors of the logistic regression coefficients for `income` and `balance` .

```{r}
bootstrap <- boot(Default, boot.fn, R = 10000)
```

## Part (d)

Comment on the estimated standard errors obtained using the `glm()` function and using your bootstrap function.

```{r}
par(mfrow = c(1, 3))
labs <- c("(Intercept)", "income", "balance")
for (i in 1:3) hist(bootstrap$t[,i], main = NA, xlab = labs[i])
```

As one can see the intercept distributions obtained by bootstrap are centred around the estimates achieved by the model fit `glm.fit`. The standard errors reported are also very similar.
