---
title: Final Project
author: Kári Hlynsson
date: April 2024
output: html_document
---

```{r include = FALSE}
library(dplyr)
library(pbapply)
library(scales)
library(latex2exp)
library(randomForest)
library(glmnet)
library(e1071)
library(class)
library(ROCR)
source("pair_plot.R")

plot_roc_curve_rocr <- function(predictions, truth, title) {
  pred <- prediction(predictions, truth)
  perf <- performance(pred, "tpr", "fpr")

  plot(
    perf,
    main = title,
    xlab = "False Positive",
    ylab = "True Positive"
  )

  abline(a = 0, b = 1, lwd = 2, lty = 2)

  auc <- performance(pred, measure = "auc")
  auc_value <- auc@y.values[[1]]
  cat("AUC: ", auc_value, "\n")

  return(list(roc = perf, auc = auc_value))
}

islrblue <- "#3333b2"
islrred <- "#990000"
islrgreen <- "#009900"
```

## Part 1: Data Cleaning / Clustering (30%)
For this section, we will use the `MAS202M_s24_synthetic_data_20240321.csv`,
which is a synthetic data set.

1. Prepare the provided data set (or another data set of your choice) for 
further analysis. Using the tools at your disposal for data cleaning, prepare 
the data set to maximize the expected performance of a clustering algorithm 
trained on the data set. Note the steps taken, whether or not they result in 
modifications to the data set. 

2. Build a model to cluster the data set. Describe the resulting model and its
performance, i.e., how well the clusters fit the data. 

### Solution

#### 1. Data cleaning

Start by reading in the data set:

```{r 11_import_data}
synthetic_dat <- read.table(
  "MAS202M_s24_synthetic_data_20240321.csv",
  sep = ",",
  head = TRUE
)
```

Summary of the data:

```{r 11_data_summary}
summary(synthetic_dat)
```

The data contains some `NA` measurements. We will have to address this and a
couple of other issues before we can move on to the clustering step. It would
be wise to start with a pair plot of the variables in the data to get a feel
for the terrain:

```{r 11_pair_plot}
pair_plot(synthetic_dat[, -2])
```

The `id` variable is an index of the data which is redundant 
information and can be removed. Make the change:

```{r 11_drop_index}
synthetic_dat <- synthetic_dat[, -1]
```

More interestingly, we see that covariates
`col3` and `col6` correspond perfectly to each other -- that is, we have total
collinearity. Let us investigate this matter further:

```{r 11_shenanigans}
synthetic_dat$col3 == synthetic_dat$col6
```

From this obnoxiously large block of output we see that the predictors are the
exact same except at some `NA` values. What might we infer about the predictors
at these points?

```{r 11_tomfoolery}
synthetic_dat |> filter(is.na(col3) | is.na(col6))
```

The data seem to be exclusively missing -- that is, when `col3` is missing,
`col 6` is not, and they are otherwise identical. This leads me to believe that
the two columns are actually one and the same but have somehow been split across
the data set (space and time, if you will). Therefore we will merge these two
columns and throw `col6` away:

```{r 11_drop_col6}
synthetic_dat <- synthetic_dat |>
  mutate(col3 = ifelse(is.na(col3), col6, col3)) |>
  select(-col6)
```

For convenience and legibility, reorder observations in increasing order of
`id` and reindex the row names:

```{r 11_reorder_data}
ord <- gsub("sample", "", synthetic_dat$id) |>
  as.numeric() |>
  order()

synthetic_dat <- synthetic_dat[ord, ]

rownames(synthetic_dat) <- 1:nrow(synthetic_dat)
```

Quite interestingly, looking at the `tail()` of these data show that a handful
observations seem to be missing `id`s, but are otherwise perfectly intact. We
will address this at a later time.

```{r 11_data_tail}
synthetic_dat |> tail(10)
```

Now we move on to cleaning duplicates and missing values from the data. Start
with the former:

```{r 11_dup_obs}
dup <- duplicated(synthetic_dat)
sum(dup)
```

A total of 97 observations are total duplicates of others. Let's remove these
from the data set. We will call the resulting output `synthetic` instead of
`synthetic_dat` to signify the cleaned data set:

```{r 11_remove_dup}
synthetic <- synthetic_dat |> distinct(.keep_all = TRUE)
```

Other duplicates are a bit more well disguised, but we start with those that
share the same `id`s and *almost* the same measurements of `col1` through
`col6` but contain `NA` measurements in one of the predictors:

```{r}
dup_ids <- which(duplicated(synthetic$id))

dup_ids_tmp <- synthetic[c(dup_ids, dup_ids - 1), ] |>
  arrange(id) |>
  filter(id != "")

dup_ids_tmp
```

Which rows do we want to drop?

```{r}
dup_ids_drop <- which((rowSums(apply(dup_ids_tmp, 2, is.na)) == TRUE)) |>
  names() |>
  as.numeric()

dup_ids_drop
```

We also have one sneaky duplicate which is missing an `id` but is otherwise
identical to another observation:

```{r}
dup_id_na <- duplicated(synthetic[, 2:6]) |> which()
synthetic |> filter(col1 == synthetic[dup_id_na, 2])
```

Remove the observations:

```{r}
synthetic <- synthetic[-dup_ids_drop, ] # 2 dropped (NA-identical)
synthetic <- synthetic[-dup_id_na, ] # drop identical obs. with no ID
```

Now that we have removed all duplicates from the data set, we can remove the
`id` column since it no longer contains any useful information.

```{r}
synthetic <- synthetic |> select(-id)
```

Let us finally impute the missing values in the data set. We can take a look
at how much of the data set they make up:

```{r}
is.na(synthetic) |> colMeans()
```

This is hardly worth worrying over. We could in fact simply remove these
observations from the data set, but we will impute them instead using the
mean of their missing covariates since they still contain some information.

```{r}
synthetic <- synthetic |> mutate(
  across(col1:col5, function(c) ifelse(is.na(c), mean(c, na.rm = TRUE), c))
)
```

Let's finish with a pair plot of the cleaned data set:

```{r}
pair_plot(synthetic)
```

And a summary for good luck:

```{r}
summary(synthetic)
```

Everything seems to be in order.

#### 2. Clustering

We will explore two clustering methods, namely hierarchical clustering and
K-means. We will also scale and transform the data in various ways and
compare the results of clustering with the respective transformed data sets.

Firstly, we transform the data using PCA to see whether dimensionality
reduction is possible and whether some discernible clustering pattern arises
when viewing the first two principal components.

Start by computing the PCA and the PVE of the various principal components.
We will do this on standardised data, since the spread and centre of the data
vary a bit across columns.

```{r}
pr_out <- prcomp(synthetic, scale = TRUE)

pr_var <- pr_out$sdev^2
pve <- pr_var / sum(pr_var)
```

These can be used to generate a biplot and a scree plot:

```{r}
biplot(
  pr_out,
  scale = 0,
  col = c(alpha(4, 0.25), 4),
  cex = c(0.5, 1.25)
)
abline(h = 0, lty = 2, col = "#cccccc")
abline(v = 0, lty = 2, col = "#cccccc")

par(mfrow = c(1, 2))
plot(
  pve,
  type = "b",
  xlab = "Principal Component",
  ylab = "PVE",
  ylim = c(0, 1)
)

plot(
  cumsum(pve),
  type = "b",
  xlab = "Princap Component",
  ylab = "Cumulative PVE",
  ylim = c(0, 1)
)
```

The biplot shows a cluster pattern separated mainly by the first principal
component. Perhaps this is the key to obtaining a sensible clustering of the
data. From the scree plots we also see that the first two components account
for roughly 60% of the variation in the data, which is quite substantial. It
would thus be sensible to consider these two axes when clustering the data.

Hierarchical clustering is performed on the data below for the raw, scaled
and PCA-transformed data. These have all been tinkered with to obtain the
(subjectively) most pleasing results.

```{r}
# Hierarchical clustering - Out of the box
hclust <- hclust(dist(synthetic), method = "average")
plot(hclust, main = "Raw data, average", labels = FALSE, sub = NA, xlab = NA)

# Pair plot of clustering
hclust_out <- cutree(hclust, 2)

pair_plot(synthetic, col = hclust_out + 2, cex = 0.2)

# Hierarchical clustering -- Scaled
hclust_sc <- hclust(dist(scale(synthetic)), method = "complete")
plot(hclust_sc,
  main = "Scaled data, complete", labels = FALSE, sub = NA,
  xlab = NA
)

# Pair plot of clustering
hclust_sc_out <- cutree(hclust_sc, 2)

pair_plot(synthetic, col = hclust_sc_out + 2, cex = 0.2)

# Hierarchical clustering -- PC1 & PC2
hclust_pca <- hclust(dist(pr_out$x[, 1:2]), method = "complete")
plot(hclust_pca,
  main = "PCA data, complete", labels = FALSE, sub = NA,
  xlab = NA
)

# Pair plot of clustering
hclust_pca_out <- cutree(hclust_pca, 2)

pair_plot(synthetic, col = hclust_pca_out + 2, cex = 0.2)

# Plot of PC1 & PC2 and clustering
plot(pr_out$x[, 1], pr_out$x[, 2],
  col = hclust_pca_out + 2, pch = 20,
  xlab = "PC1", ylab = "PC2"
)
```

Hierarchical clustering seems to yield the most sane results on the raw 
data with completed linkage. The clustering on the first two principal
components seems to violate our expectations a bit -- let's see whether this
issue is less prominent with K-means.

```{r}
# K Means -- Out of the box performance
kmeans <- kmeans(synthetic, centers = 2, nstart = 50)
kmeans$cluster <- ifelse(kmeans$cluster == 1, 2, 1)
pair_plot(synthetic, col = kmeans$cluster + 2, cex = 0.2)

# K Means -- Scaled data
kmeans_sc <- kmeans(scale(synthetic), centers = 2, nstart = 50)
kmeans_sc$cluster <- ifelse(kmeans_sc$cluster == 1, 2, 1)
pair_plot(synthetic, col = kmeans_sc$cluster + 2, cex = 0.2)

# K Means -- PC1 & PC2
kmeans_pr <- kmeans(pr_out$x[, 1:2], centers = 2, nstart = 50)
kmeans_pr$cluster <- ifelse(kmeans_pr$cluster == 1, 2, 1)
pair_plot(synthetic, col = kmeans_pr$cluster + 2, cex = 0.2)

# Plot of PC1 & PC2 and clustering
plot(pr_out$x[, 1], pr_out$x[, 2],
  col = kmeans_pr$cluster + 2, pch = 20,
  xlab = "PC1", ylab = "PC2"
)
```

K-means looks fantastic on the first two principal components and is very
much in line with what one would expect of the data. 

## Part 2: Prediction / Inference (70%)

For this section, we provide a data set derived from proteomics measurements 
performed by Marie-Pierre Dube. The file `proteomics_dube_2023.csv` contains 
plasma protein levels for 1,462 proteins (columns) for 40 samples (rows); 
10 subjects, each measured at two acclimation stages and in two thermal states.

1. Pick a quantitative variable in the data set. Build a prediction model for 
the variable using two different methods. At least one of the prediction models
should be able to handle non-linear relationships between dimensions. Tune the
parameters of each model and compare their performances. discuss the outcome,
including the strengths and weaknesses of each type of model relative to the 
data set in question. what (if anything) can you infer about the structure of 
the data from the models?

2. Pick a categorical variable in the data set (or create one by thresholding 
a continuous variable). Build a classifier for the variable using two different
methods. At least one of the classifiers should be able to fit non-linear 
decision boundaries. Tune the parameters of each classifier and compare their 
performances. Discuss the outcome, including the strengths and weaknesses of 
each type of classifier relative to the data set in question. What 
(if anything) can you infer about the structure of the data from the models?


## Solution

### 1. Quantitative variable

I will investigate the `EHBP1` variable. We start by importing the data set.
The `SubjectID` column is removed since it does not have any relevance in the
analyses we are about to perform.

```{r 2_data_in}
# read in the data set
dube <- read.csv("proteomics_dube_2023.csv", stringsAsFactors = TRUE)[, -1]
n <- nrow(dube)
p <- ncol(dube)
```

We will use both cross-validation during the fitting process, but we also want
to have a measure of how well our models perform on unseen data. Thus we must
perform a train-test split:

```{r 2a_data_split}
# set RNG seed for reproducibility
set.seed(42)

# perform train-test split
prop_test <- 0.75
s <- sample(1:n, 0.75 * n)
train <- dube[s, ]
test <- dube[setdiff(1:n, s), ]
```

For the first variable, `EHBP1`, we will employ two methods: LASSO and random
forests. We start with the former.

#### LASSO

We fit LASSO to the training data using the `cv.glmnet` command, which performs
\(k\)-fold cross-validation in order to determine the optimal value of the
regularisation parameter, \(\lambda\). We will perform LOOCV, as shown below:

```{r 2a_LASSO_fit, warning = FALSE}
# arrange the data into model matrix format so it is suitable for cv.glmnet
x_train <- model.matrix(EHBP1 ~ ., data = train)[, -1]
x_test <- model.matrix(EHBP1 ~ ., data = test)[, -1]
y_train <- train[, "EHBP1"]
y_test <- test[, "EHBP1"]

# perform the cross-validation
cv_lasso <- cv.glmnet(
  x_train,
  y_train,
  nfolds = 30,
  thresh = 1e-12
)

# select the optimal regularisation parameter
lambda_opt <- cv_lasso$lambda.min
```

The optimal parameter is \(\lambda = `r lambda_opt`\).

```{r 2a_LASSO_plot}
# generate a plot of model coefficients as a fn. of regularisation param.
# and the LOOCV error as a fn. of
par(mfrow = c(1, 2))
plot(
  log(cv_lasso$lambda),
  cv_lasso$cvm,
  type = "l",
  xlab = expression(log(lambda)),
  ylab = "LOOCV error",
  ylim = c(0, 1)
)

lines(log(cv_lasso$lambda), cv_lasso$cvup, lty = 2)
lines(log(cv_lasso$lambda), cv_lasso$cvlo, lty = 2)
abline(v = log(lambda_opt), lty = 2)

plot(
  cv_lasso$glmnet.fit,
  "lambda",
  xlab = expression(lambda),
  ylab = "Coefficients"
)

abline(v = log(lambda_opt), lty = 2)
```

Let us take a look at the coefficients yielded by using the optimal 
\(\lambda\):

```{r}
# find and plot non-zero model coefficients
coefs_lasso <- predict(cv_lasso, s = lambda_opt, type = "coefficients")
coefs_lasso_nz <- coefs_lasso[coefs_lasso[, 1] != 0, ]

coefs_lasso_nz |>
  sort(decreasing = TRUE) |>
  barplot(names.arg = names(coefs_lasso_nz), cex.names = 0.75, col = NA)
```

The LASSO model reduces the amount of covariates involved in 
predicting the response from 1463 down to 8, excluding the intercept. One of 
the advantages of using LASSO is its capability to shrink coefficient estimates
to zero using the regularisation term in its cost function which is of the form
\(\lambda\sum_{j = 1}^p |\beta_j|\), where \(\beta_j\) is the effect size for 
the \(j\)th covariate. Thus it is a reasonable choice given the high
dimensional setting posed by the data considered. However, using this model 
assumes a linear relationship between the causal covariates and the response,
which may be overly stringent.

In light of this, let us measure the performance of the optimal model:

```{r}
# lasso test set mse
lasso_pred <- predict(cv_lasso, s = lambda_opt, newx = x_test)
mse_lasso <- mean((lasso_pred - y_test)^2)
mse_lasso
```

The MSE is \(`r mse_lasso`\), which is not bad at all. The fit can also be
shown visually as below:

```{r}
# identity plot of data vs. predictions
plot(
  lasso_pred,
  y_test,
  xlab = "Predicted response",
  ylab = "Observed repsonse"
)
abline(a = 0, b = 1, lty = 2)
```

The model tends to overshoot a bit in the left extreme. It is interesting to
view the non-zero covariates on a pair plot with our reponse to see whether
there is any discernible relationship:

```{r}
# pair plot of non-zero coefficients with response
names_lasso <- c("EHBP1", names(coefs_lasso_nz)[-1])
pair_plot(dube[, names_lasso], cex = 0.2)
```

The `EHBP1` column shows quite strong linear relationships between the
covariates and the response, which we can also see by viewing the row which
`EHBP1` belongs to. There is also a very strong correlation between the
covariates themselves, which may suggest that these proteins might be linked in
some way, such as coming from the same amino-acid sequences but undergoing
post-translational modifications to assume different functionalities. However,
a lot more testing would have to take place until one could be sure.

#### Random forest

Now let us try using a random forest model to describe the relationship. The
optimal value of `mtry` is first determined by measuring the test MSE of the
fitted models at each turn:

```{r message = FALSE}
# fit and tune a random forest regressor
mtry <- c(seq(10, p - 1, by = 100)) |> sort()

rf_fits <- pblapply(mtry, function(mt) {
  rf_fit <- randomForest(
    EHBP1 ~ .,
    data = train,
    xtest = x_test |> data.frame(),
    ytest = y_test,
    mtry = mt,
    ntree = 500,
    importance = TRUE
  )
})

rf_perf <- pblapply(rf_fits, function(fit) {
  list(mse = fit$test$mse, balanced = fit$test$mse[500])
})
```

The `rf_fits` list contains the fitted models whereas `rf_perf` contains the
performances. `rf_perf` consists of two atomic vectors, `mse` and `balanced`,
where the first gives the CV error for different tree sizes, whereas the
latter gives the CV error at `ntree` = 500.

The best model is found and the CV error is plotted as a function of tree size:

```{r}
# find and plot best model
best_model <- sapply(rf_perf, function(i) i$balanced) |> which.min()

plot(
  1,
  type = "n",
  xlab = "Number of trees",
  ylab = "Test set MSE",
  xlim = c(1, 500),
  ylim = c(0, 1)
)

for (i in 1:length(rf_perf)) {
  if (i != best_model) {
    lines(
      rf_perf[[i]]$mse,
      col = ifelse(best_model == i, "red", alpha("black", 0.25)),
      lwd = ifelse(best_model == i, 3.0, 1.0),
      type = "l"
    )
  }
}

lines(rf_perf[[best_model]]$mse,
  lwd = 4,
  col = 4,
  type = "l",
  ylim = range(rf_perf),
  xlab = "No. trees",
  ylab = "CV error"
)


legend(
  "topright",
  inset = c(0.01, 0.01),
  bty = "n",
  legend = c(paste0("mtry = ", mtry[[best_model]])),
  col = 4,
  lwd = 3.5
)
```

The value of `mtry` which minimises the CV error turns out to be 
`r mtry[[best_model]]`. Shown below are variable importance plots for
the predictors involved in the random forest model, some of which we recognise
from out fitting of the lasso model:

```{r}
# variable importance plot
varImpPlot(rf_fits[[best_model]], main = NA)

# get 30 highest scoring predictors in terms of purity and mse
inc_pur <- importance(rf_fits[[best_model]]) |>
  data.frame() |>
  arrange(IncNodePurity) |>
  tail(30) |>
  rownames()

inc_mse <- importance(rf_fits[[best_model]]) |>
  data.frame() |>
  arrange(X.IncMSE) |>
  tail(30) |>
  rownames()
```

As before we can visually validate the relationships that are specified by the
random forest model. I will show some pair plots for a subsets of the variables
in `inc_mse` for brevity:

```{r}
# pair plot of a sample of importance predictors in random forest
pair_plot(dube[, c("EHBP1", inc_mse |> sample(8))], cex = 0.2)
```

The result is quite interesting. Few if not none of these variables held any 
weight in the LASSO regression, but many of them seem to be quite strongly 
linearly correlated with the response, as is evidenced by the plot produced 
above. One of the benefits of random forest is its capacity to work in a
high-dimensional such as the one consider here where \(n \ll p\). It is a
powerful method due to its capacity to model non-linear boundaries but
provides robust predictive capabilities since it decorrelates the decision
trees used to form the ensemble vote by both bootstrapping the training data
set and picking a subset of predictors that are considered when constructing
new trees. Here we see that it is able to pick up on effects from covariates
that would go unnoticed in a traditional regularised linear regression such as
when using Ridge or LASSO, which may be very useful when doing a broad sweep of
the data when one is searching for potential biomarkers and mostly focused on
pruning the data.

How does the random forest model perform on the test data set?

```{r}
rf_fits[[best_model]]$test$mse[500]
```

Random forest regression performs similarly to the LASSO model we considered
above.

```{r}
plot(
  rf_fits[[best_model]]$test$predicted,
  y_test,
  xlab = "Predicted response",
  ylab = "Observed response"
)
abline(a = 0, b = 1, lty = 2)
```

It looks good, but tends to overshoot at the left extreme, similarly as we
saw in the LASSO model.

### 2. Categorical variable

We will investigate the categorical variable `temp` to see if any protein
measurements along with the acclimation state can be used to meaningfully
predict whether an individual is of normal body temperature (*normothermic*)
or whether he is hot (*hyperthermic*).

In order to reduce the number of involved covariates we will transform the
data using PCA:

```{r}
# calculate PCA and create scree plot
pr_out <- prcomp(dube[, 3:p], scale = TRUE)

pr_var <- pr_out$sdev^2
pve <- pr_var / sum(pr_var)

par(mfrow = c(1, 2))
plot(
  pve,
  xlab = "Number of Principal Components",
  ylab = "PVE",
  type = "b"
)
abline(v = 15, lty = 2, col = "red")
plot(
  cumsum(pve),
  xlab = "Number of Principal Components",
  ylab = "Cumulative PVE",
  type = "b"
)
abline(v = 15, lty = 2, col = "red")
```

This is beneficial in reducing the computation time many of the model fittings
done here below, while also retaining a lot of the variability in the original
data set. From the scree plot and after some tinkering around with results I 
determined that 15 principal components would be suitable for fitting the 
models.

Let us now generate a train-test split of the PCA-transformed data, this time
a 50% split.

```{r}
# train test split for classifiers
set.seed(30)
s <- sample(1:n, 0.5 * n, replace = FALSE)
dube_pr <- cbind(dube[, 1:2], pr_out$x[, 1:15])
train_pr <- dube_pr[s, ]
test_pr <- dube_pr[setdiff(1:n, s), ]
```

#### SVM

In this step we will fit SVMs with different kernels and optimise for each
of the possible parameters. SVMs are powerful in the high dimensional setting
since fitting a \(p - 1\) hyperplane to separate the data relies and have the
desirable quality of mapping high-dimensional data to dimensions of lower
dimension while retaining the information, often in such a way that a linear
decision boundary can be identified. However, they can be very difficult to
interpret.

We start with the linear kernel, or in other words a support vector classifier,
where the optimisation is over the parameter `cost`:

```{r}
# fit SVC
svm_fit_pr_lin <- tune(
  svm,
  temp ~ .,
  data = train_pr,
  kernel = "linear",
  ranges = list(
    cost = 10^seq(-3, 3, length.out = 20)
  )
)

lin_perf <- svm_fit_pr_lin$best.performance
lin_param <- svm_fit_pr_lin$best.parameters
lin_pred <- predict(svm_fit_pr_lin$best.model, train_pr)

lin_perf
lin_param

plot_roc_curve_rocr(
  as.numeric(lin_pred) - 1,
  as.numeric(test_pr$temp) - 1,
  "SVC ROC curve"
)
```

The optimal performance is `r lin_perf` with `cost` \(= `r lin_param$cost`\).
Let's evaluate this model on the test set:

```{r}
# SVC performance
svm_lin_opt <- summary(svm_fit_pr_lin)$best.model
svm_lin_pred <- predict(svm_lin_opt, test_pr)
lin_cfmat <- table(svm_lin_pred, test_pr$temp)
lin_cfmat
```

The SVC correctly identifies observations in the test set in
\(`r (lin_cfmat[1, 1] + lin_cfmat[2, 2]) / sum(lin_cfmat) * 100`\%\) of cases.

Next we consider a SVM with a polynomial kernel, which is optimised over the
parameters `cost` and `degree`:

```{r}
# Polynomial kernel SVM
svm_fit_pr_poly <- tune(
  svm,
  temp ~ .,
  data = train_pr,
  kernel = "polynomial",
  ranges = list(
    cost = 10^seq(-3, 3, length.out = 20),
    degree = 1:8
  )
)

poly_perf <- svm_fit_pr_poly$best.performance
poly_param <- svm_fit_pr_poly$best.parameters
poly_pred <- predict(svm_fit_pr_poly$best.model, test_pr)

plot_roc_curve_rocr(
  as.numeric(poly_pred) - 1,
  as.numeric(test_pr$temp) - 1,
  "SVM polynomial kernel ROC curve"
)
```

The optimal performance is `r poly_perf` with the parameters
`cost` \(= `r poly_param$cost`\) and `degree` \(= `r poly_param$degree`\).
It is interesting to note that the kernel specified is in some sense similar
to a linear kernel since the degree of the polynomial involved is of degree one.
Model performance on the test set is evaluated below:

```{r}
# Polynomial kernel SVM performance
svm_poly_opt <- summary(svm_fit_pr_poly)$best.model
svm_poly_pred <- predict(svm_poly_opt, test_pr)
poly_cfmat <- table(svm_poly_pred, test_pr$temp)
poly_cfmat
```

The SVM with a polynomial kernel identifies test set classes in
\(`r (poly_cfmat[1, 1] + poly_cfmat[2, 2]) / sum(poly_cfmat) * 100`\%\) of
cases.

Lastly we consider a radial kernel, where we optimise over parameters `cost`
and `gamma` (\(\gamma\)).

```{r}
# Radial kernel SVM
svm_fit_pr_radial <- tune(
  svm,
  temp ~ .,
  data = train_pr,
  kernel = "radial",
  ranges = list(
    gamma = 10^seq(-3, 2, length.out = 20),
    cost = 10^seq(-3, 3, length.out = 20)
  )
)

radial_perf <- svm_fit_pr_radial$best.performance
radial_params <- svm_fit_pr_radial$best.parameters
radial_pred <- predict(svm_fit_pr_radial$best.model, test_pr)

plot_roc_curve_rocr(
  as.numeric(radial_pred) - 1,
  as.numeric(test_pr$temp) - 1,
  "SVM radial kernel ROC curve"
)
```

The best performance is \(`r radial_perf`\) with parameters `cost` 
\(= `r radial_params$cost`\) and \(\gamma = `r radial_params$gamma`\).
Let's take a look at prediction accuracy:

```{r}
# Radial kernel SVM performance
svm_radial_opt <- summary(svm_fit_pr_radial)$best.model
svm_radial_pred <- predict(svm_radial_opt, test_pr)
radial_cfmat <- table(svm_radial_pred, test_pr$temp)
radial_cfmat
```

The SVM with a radial kernel identifies test set classes in
\(`r (radial_cfmat[1, 1] + radial_cfmat[2, 2]) / sum(radial_cfmat) * 100`\%\) of
cases.

We have thus seen that using SVM, we obtain a maximum accuracy of \(70\%\) with
the linear and radial kernel.

#### KNN

We will finish this project by considering KNN as a classifier. This powerful
method has been known to suffer from some problems when employed in a 
high-dimensional setting, but we will see that this is not the case for these 
data. Tuning the model is a simple matter:

```{r}
# KNN
knn_tune <- tune.knn(
  train_pr[, -c(1, 2)],
  train_pr[, 1],
  k = 1:16,
  tunecontrol = tune.control(
    nrepeat = 10,
    cross = 5
  )
)

knn_perf <- knn_tune$best.performance
knn_param <- knn_tune$best.parameters
```

The best performance from the tuning process is `r knn_perf` with \(k = 
`r knn_param$k`\).

Produce a plot of the CV error as a function of \(k\):

```{r}
# plot of CV error as a fn. of K
plot(knn_tune$performances$k, knn_tune$performances$error,
  type = "b",
  xlab = "K", ylab = "CV error"
)
```

We will now access the optimal model and measure performance on the test set:

```{r}
# Optimal KNN model
knn_opt <- knn(train_pr[, -c(1, 2)], test_pr[, -c(1, 2)], train_pr[, 1], k = 9)
knn_cfmat <- table(knn_opt, test_pr$temp)
knn_cfmat
```

Produce a ROC curve of KNN:

```{r}
# ROC curve for KNN
plot_roc_curve_rocr(
  as.numeric(knn_opt) - 1,
  as.numeric(test_pr$temp) - 1,
  "KNN ROC curve"
)
```

KNN with \(k = `r knn_param$k`\) has an accuracy of 
\(`r sum(diag(knn_cfmat)) / sum(knn_cfmat) * 100`\%\), which is the best 
performance so far. KNN is a strong out-of-the-box classifier and is 
advantageous in that it is easy to tune, is non-parametric and very intuitive,
which for examples the SVMs are sometimes not.
