---
title: Exercise 9.4
date: March 2024
output: html_document 
author: Kári Hlynsson
---

Generate a simulated two-class data set with 100 observations and two features
in which there is a visible but non-linear separation between the two classes.
Show that in this setting, a support vector machine with a polynomial kernel
(with degree greater than 1) or a radial kernel will outperform a support
vector classifier on the training data. Which technique performs best on the
test data? Make plots and report training and test error rates in order to back
up your assertions.

The below code shows the data generation process. We start by generating 100
\(x\)-values where \(x_i \sim \mathcal N(3.5, 1.75^2)\) and generate an
\(y\)-coordinate using \(y_i \sim \mathcal N(x_i^2 + 2x_i + 4, \epsilon_i)\)
where \(\epsilon_i \sim \mathcal N(0, 20^2)\). We classify the \(y\)-values
into two classes, `blue` and `orange`, using the function
\[
  \chi_y(x) = 0.35 e^x
\]
such that \(c_i = \) `blue` if \(y_i > \chi_y(x_i)\), and `orange` otherwise,
where \(c_i\) denoted the class label of the \(i\)th observation. The data
points along with the classification boundary are shown in the figure below.

```{r}
set.seed(1)

p <- function(x) 0.35 * exp(x)
x <- rnorm(100, mean = 3.5, sd = 1.75)
y <- x^2 + 2 * x + 4 + rnorm(100, mean = 0, sd = 20)
k <- c()

for (i in 1:100) {
  k[i] <- ifelse(y[i] > p(x[i]), "blue", "orange")
}

plot(function(x) 0.35 * exp(x), from = -3, to = 8,
     xlim = range(x), ylim = range(y), lwd = 2.5,
     ylab = "y")
points(x, y, col = k, pch = 20, cex = 1.25)
abline(h = 0, lty = 2)
abline(v = 0, lty = 2)

# train / test split
n_train <- 75
df <- data.frame(x = x, y = y, k = as.factor(k))
tr <- sample(1:100, n_train)
ts <- setdiff(1:100, tr)

train <- df[tr, ]
test <- df[ts, ]
```

Now let us try to classify these data using support vector classifiers and
support vector machines and compare the two methods.


## SVC

```{r}
library(e1071)
library(caret)
library(scales)

svcfit <- svm(k ~ y + x, data = train, cost = 10, kernel = "linear")
svcfit_pred <- predict(svcfit, test)

svcfit_conf <- confusionMatrix(svcfit_pred, test$k)
svcfit_conf
```

The accuracy with an out-of-the-box SVC is 0.8 -- mediocre. Visualise the
decision boundary:

```{r}
plot(
  svcfit,
  train,
  symbolPalette = c("blue", "orange"),
  svSymbol = 10,
  dataSymbol = 20,
  color.palette = function(x) c(alpha("blue", 0.7), alpha("orange", 0.7))
)
```

CV over cost parameter to obtain more accurate model:

```{r}
svcfit_tuned <- tune(
  svm,
  k ~ y + x,
  data = train,
  kernel = "linear",
  ranges = list(
    cost = 10^seq(-2, 2, length.out = 20)
  )
)

tune_perf <- svcfit_tuned$performances

plot(tune_perf$cost, tune_perf$error, type = "b",
     xlab = "Cost", ylab = "Error")
```

The best value of cost is

```{r}
svcfit_tuned$best.parameters
```

Recover the best model:

```{r}
svcfit_opt <- svcfit_tuned$best.model
svcfit_opt_pred <- predict(svcfit_opt, test)

plot(
  svcfit_opt,
  train,
  symbolPalette = c("blue", "orange"),
  svSymbol = 10,
  dataSymbol = 20,
  color.palette = function(x) c(alpha("blue", 0.7), alpha("orange", 0.7))
)

svcfit_opt_conf <- confusionMatrix(svcfit_opt_pred, test$k)
svcfit_opt_conf
```

## SVM

```{r}
svmfit_poly <- svm(k ~ y + x, data = train, kernel = "polynomial", d = 3,
                   cost = 10)

plot(
  svmfit_poly,
  train,
  symbolPalette = c("blue", "orange"),
  svSymbol = 10,
  dataSymbol = 20,
  color.palette = function(x) c(alpha("blue", 0.7), alpha("orange", 0.7))
)
```

Let's try optimising the polynomial kernel SVM:

```{r}
svmfit_poly_tune <- tune(
  svm,
  k ~ y + x,
  data = train,
  kernel = "polynomial",
  ranges = list(
    d = 1:8,
    cost = c(0.001, 0.01, 0.1, 1, 10, 100, 1000)
  )
)

svmfit_poly_tune$best.parameters
```

Recover best model:

```{r}
svmfit_poly_opt <- svmfit_poly_tune$best.model
svmfit_poly_opt_pred <- predict(svmfit_poly_opt, test)

svmfit_poly_opt_conf <- confusionMatrix(svmfit_poly_opt_pred, test$k)

plot(
  svmfit_poly_opt,
  train,
  symbolPalette = c("blue", "orange"),
  svSymbol = 10,
  dataSymbol = 20,
  color.palette = function(x) c(alpha("blue", 0.7), alpha("orange", 0.7))
)

svmfit_poly_opt_conf
```

It is considerably more accurate 
SVM with radial basis kernel:

```{r}
svmfit_radial <- svm(k ~ y + x, data = train, kernel = "radial",
                     gamma = 0.5, cost = 10)

plot(
  svmfit_radial,
  train,
  symbolPalette = c("blue", "orange"),
  svSymbol = 10,
  dataSymbol = 20,
  color.palette = function(x) c(alpha("blue", 0.7), alpha("orange", 0.7))
)

svmfit_radial_pred <- predict(svmfit_radial, test)
svmfit_radial_conf <- confusionMatrix(svmfit_radial_pred, test$k)

svmfit_radial_conf
```

Tune the model:

```{r}
svmfit_radial_tune <- tune(
  svm,
  k ~ y + x,
  data = train,
  kernel = "radial",
  ranges = list(
    gamma = 10^seq(-2, 2, length.out = 20),
    cost = 10^seq(-2, 2, length.out = 20)
  )
)

svmfit_radial_tune$best.parameters
```

The best model:

```{r}
svmfit_radial_opt <- svmfit_radial_tune$best.model
svmfit_radial_opt_pred <- predict(svmfit_radial_opt, test)
svmfit_radial_opt_conf <- confusionMatrix(svmfit_radial_opt_pred, test$k)
svmfit_radial_opt_conf
```

The accuracy is very high.

```{r}
plot(
  svmfit_radial_opt,
  train,
  symbolPalette = c("blue", "orange"),
  svSymbol = 10,
  dataSymbol = 20,
  color.palette = function(x) c(alpha("blue", 0.7), alpha("orange", 0.7))
)
```

Radial basis kernel SVM for the win.
