---
title: Exercise 9.7
date: April 2024
output: html_document
author: Kári Hlynsson
---

In this problem, you will use support vector approaches in order to predict
whether a given car gets high or low gas mileage based on the `Auto` data set.

## Part (a)
Create a binary varable that takes on a 1 for cars with gas mileage above the
median, and a 0 for cars with gas mileage below the median.

```{r}
library(ISLR2)
data(Auto)

set.seed(1)

Auto$mpg01 <- as.factor(ifelse(Auto$mpg > median(Auto$mpg), "Above", "Below"))
Auto <- subset(Auto, select = -name)
```

## Part (b)
Fit a support vector classifier to the data with various values of `cost`, in
order to predict whether a car gets high or low gas mileage. Report the
cross-validation errors associated with different values for this parameter.
Comment on your results. Note you will need to fit the classifier without the
gas mileage variable to produce sensible results.

```{r}
library(e1071)

svc <- tune(
  svm,
  mpg01 ~ .,
  data = Auto,
  kernel = "linear",
  ranges = list(
    cost = c(0.001, 0.01, 0.1, 1, 10, 100, 1000)
  )
)
```

Let's take a look at CV error as a function of log `cost` during tuning:

```{r}
plot(log(svc$performances$cost, base = 10),
     svc$performances$error, type = "b",
     xlab = "Log(Cost)",
     ylab = "CV Error")
```

The best value for `cost` is `r svc[["best.parameters"]]`, with a CV error
rate of `r svc[["best.performance"]]`, which corresponds to a training set
accuracy of `r 1 - svc[["best.performance"]]`, which is pretty good.

## Part (c)
Now repeat (b), this time using SVMs with radial and polynomial basis kernels,
with defferent values of `gamma` and `degree` and `cost`. Comment on your
results.

Let's start with the radial kernel.

```{r}
svm_radial <- tune(
  svm,
  mpg01 ~ .,
  data = Auto,
  kernel = "radial",
  ranges = list(
    gamma = c(0.5, 1, 2, 3, 4, 5),
    cost = c(0.001, 0.01, 0.1, 1, 10, 100, 1000)
  )
)
```

Best performance is for \(\gamma = `r svm_radial$best.parameters["gamma"]`\)
and `cost` \(= `r svm_radial$best.parameters["cost"]`\). This yields a CV
error of \(`r svm_radial$best.performance`\), which corresponds to an accuracy
of \(`r 1 - svm_radial$best.performance`\).

Let's do the same for the polynomial basis kernel. We will study polynomials
up to degree six.

```{r}
svm_poly <- tune(
  svm,
  mpg01 ~ .,
  data = Auto,
  kernel = "polynomial",
  ranges = list(
    d = 1:6,
    cost = c(0.001, 0.01, 0.1, 1, 10, 100, 1000)
  )
)
```

Best performance is given by \(d = `r svm_poly$best.parameters["d"]`\) and
`cost` \(= `r svm_poly$best.parameters["cost"]`\), yielding a CV error of
\(`r svm_poly$best.performance`\), again corresponding to an accuracy of
\(`r 1 - svm_poly$best.performance`\).

## Part (d)
Make some plots to back up your assertions in (b) and (c).

We will create plots of all the decision boundaries for a given method to see
how well the method is capturing the relationships in the data. This
corresponds to 8 plots per method.

```{r}
vars <- colnames(Auto)

plot_vars <- function(fit) {
  for (var in vars[!(vars %in% c("mpg", "mpg01"))]) {
    plot(fit$best.model, Auto, as.formula(paste0("mpg ~ ", var)),
         svSymbol = 10, dataSymbol = 20)
  }
}
```

### SVC

```{r}
plot_vars(svc)
```

### SVM -- Radial

```{r}
plot_vars(svm_radial)
```

### SVM -- Polynomial

```{r}
plot_vars(svm_poly)
```
