---
title: Exercise 9.5
date: March 2024
output: html_document
author: Kári Hlynsson
---

We have seen that we can fit a SVM with a non-linear kernel in order to perform
classification using a non-linear decision boundary. We will now see that we
can also obtain a non-linear decision boundary by performing logistic
regression using non-linear transformation of the features.

## Part (a)
Generate a data set with \(n = 500\) and \(p = 2\), such that the observations
belong to two classes with a quadratic decision boundary between them. For
instance, you can do this as follows:

```
> x1 <- runif(500) - 0.5
> x2 <- runif(500) - 0.5
> y <- 1 * (x1^2 - x2^2 > 0)
```

```{r}
set.seed(123456)
x1 <- runif(500) - 0.5
x2 <- runif(500) - 0.5
y <- 1 * (x1^2 - x2^2 > 0)

df <- data.frame(x1 = x1, x2 = x2, y = y)
```

## Part (b)
Plot the observations, coloured according to their class labels. Your plot
should display \(X_1\) on the \(x\)-axis, and \(X_2\) on the \(y\)-axis.

```{r}
plot(x1, x2, col = ifelse(y == 1, "blue", "red"), xlab = expression(X[1]),
     ylab = expression(X[2]))
```

## Part (c)
Fit a logistic regression model to the data, using \(X_1\) and \(X_2\) as
predictors.

```{r}
logreg <- glm(y ~ ., data = df, family = "binomial")
```

## Part (d)
Apply this model to the *training data* in order to obtain a predicted class
label for each training observation. Plot the observations, coloured according
to the *predicted* class labels. The decision boundary should be linear.

```{r}
logreg_prob <- predict(logreg, df, type = "response")
logreg_pred <- ifelse(logreg_prob > 0.5, 1, 0)

plot(x1, x2, col = ifelse(logreg_pred == 1, "blue", "red"),
     xlab = expression(X[1]), ylab = expression(X[2]))
```

## Part (e)
Now fit a logistic regression model to the data using non-linear functions of
\(X_1\) and \(X_2\) as predictors (e.g. \(X_1^2\), \(X_1 \cdot X_2\) etc.).

```{r warning = FALSE}
logreg_mixed <- glm(y ~ x1 + I(x1^2) + x2 + I(x2^2) + I(x1 * x2), data = df,
                    family = "binomial")
```


## Part (f)
Apply this model to the *training data* in order to obtain a predicted class
label for each training observation. Plot the observations, coloured according
to the *predicted* class labels. The decision boundary should be obviously
non-linear. If it is not, then repeat (a)-(e) until you come up with an example
in which the predicted class labels are obviously non-linear.

```{r}
logreg_mixed_prob <- predict(logreg_mixed, type = "response")
logreg_mixed_pred <- ifelse(logreg_mixed_prob > 0.5, 1, 0)

plot(x1, x2, col = ifelse(logreg_mixed_pred == 1, "blue", "red"))
```

Captures the true decision boundary quite well.

## Part (g)
Fit a support vector classifier to the data with \(X_1\) and \(X_2\) as
predictors. Obtain a class prediction for each training observation. Plot the
observations, coloured according to the *predicted class labels*.

```{r}
library(e1071)

svc_tune <- tune(
  svm,
  as.factor(y) ~ x1 + x2,
  data = df,
  kernel = "linear",
  ranges = list(
    cost = c(0.01, 0.1, 1, 10, 100, 500)
  )
)
```

```{r}
svc_fit <- svc_tune$best.model
svc_pred <- predict(svc_fit, df)

plot(x1, x2, col = ifelse(svc_pred == 1, "blue", "red"))
```

Absolutely terrible.

## Part (h)
Fit a SVM using a non-linear kernel to the data. Obtain a class prediction for
each training observation. Plot the observations, coloured according to the
*predicted class labels*.

```{r}
svm_tune <- tune(
  svm,
  as.factor(y) ~ x1 + x2,
  data = df,
  kernel = "radial",
  ranges = list(
    gamma = 10^seq(-2, 3, length.out = 10),
    cost = 10^seq(-2, 3, length.out = 10)
  )
)
```

Analyse

```{r}
svm_fit <- svm_tune$best.model
svm_pred <- predict(svm_fit, df)

plot(x1, x2, col = ifelse(svm_pred == 1, "blue", "red"))
```

## Part (i)
Comment on your results.

The linear logistic regression model and the support vector classifier both
failed to recognise the decision boundary, whereas the mixed term logistic
regression model and the radial kernel support vector machine models did so
very well. The SVM approach required less manual work than the logistic
regression with mixed terms.
