---
title: "Week 1 — Exercise 2.8"
author: "Kári Hlynsson"
output: 
  bookdown::html_document2:
    toc: true
    toc_float:
      toc_collapsed: false    
      smooth_scroll: false
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")

panel.hist <- function(x, ...)
{
    usr <- par("usr")
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, ...)
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y)
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    text(0.5, 0.5, txt, cex = 1.5)
}

pair_plot <- function(x) { pairs(x, diag.panel = panel.hist, upper.panel = panel.cor) }
```

# Packages

```{r message = FALSE}
library(ISLR2)
library(tidyverse)
library(magrittr)
library(GGally)
library(gridExtra)

theme_set(theme_classic())
```


# Description of data set

Contained in this document is my solution of Exercise 2.8 from *An Introduction to Statistical Learning*. In this exercise, the data set `College` is used, which contains a number of variables for 777 different universities and colleges in the US. The variables in the data as well as a short description is shown below in Table \@ref(tab:data-info).

```{r data-info, echo = FALSE}
tab_data_info <- read_delim("ex2.8-data-info", show_col_types = FALSE)
knitr::kable(tab_data_info, caption = "Description of the `College` data set.")
```

# Exploring the data

Start by importing the data set using the `data` command. This is possible since the `ISLR2` package has been installed. The row names in the data set are already configured to be the names of the universities and colleges, so there is no need to make changes to that part.

```{r}
# Load the data set
data(College)
attach(College)

 # Look at data types
glimpse(College)
```

The data set consists of 777 observations with 18 features per observation. All the variables are numeric except for `Private`, which is a factor (binary) variable. There is no change necessary here. Let us summarise the variables:

```{r}
summary(College)
```

```{r message = FALSE}
pair_plot(College[,1:10])
```

Side by side boxplots of `Outstate` versus `Private`:

```{r}
boxplot(Outstate ~ Private, ylab = "Out-of-state tuition (USD)")
```

Create the `Elite` variable:

```{r}
College %<>% mutate(Elite = factor(ifelse(College$Top10perc > 50, "Yes", "No")))
```

Let's look at how many of the universities fall under the elite category:

```{r}
College %>% count(Elite)
```

78 of the 777 colleges and universities, or about 10% fall under the elite category. This is quite satisfying. We might also be interested in seeing which of the universities have the highest percentage of elite students from their respective high schools:

```{r}
College %>% 
  filter(Elite == "Yes") %>%
  arrange(desc(Top10perc)) %>%
  select(Private, Apps, Accept, Enroll, Top10perc) %>%
  head(10) %>%
  knitr::kable(caption = "Top 10 universities in terms of elite student percentage.")
```

These names are certainly not unfamiliar, at least not to an undergraduate student.
Let us next create some histograms of a selection of the quantitative variables.

```{r message = FALSE, fig.cap = "Histograms of quantitative variables in the data set."}
cols <- College |> select_if(is.numeric) |> colnames()

par(mfrow = c(3, 4))
for (c in cols) hist(College[,c], main = NA, xlab = paste(c))  
```

Does tuition affect the rate of graduation?

```{r}
plot(x = Outstate, y = Grad.Rate)
abline(lm(Grad.Rate ~ Outstate), col = "red", lty = 2, lwd = 1.5)
```

It would seem that higher tuition correlates with a higher graduation rate. However notice that we have a value of graduation rate which exceeds 100%,
which is not possible. Let's fix this and then reproduce the plot.
```{r}
Grad.Rate <- ifelse(Grad.Rate >= 100, 100, Grad.Rate)
plot(x = Outstate, y = Grad.Rate)
abline(lm(Grad.Rate ~ Outstate), col = "red", lty = 2, lwd = 1.5)
```


```{r}
lin_grad <- lm(Grad.Rate ~ Outstate)
summary(lin_grad)
```

Indeed, the tuition seems to have some bearing on the graduation rate. Maybe this model can be improved by looking at
the quality of education the institution provides.

```{r}
lin_grad2 <- update(lin_grad, . ~ . + Elite)
summary(lin_grad2)
```

Indeed the quality of students entering the school seems to be a significant predictor for rate of graduation. This is not surprising! Let's take a look:

```{r}
plot(Grad.Rate ~ Elite)
```

```{r}
plot(Outstate ~ Elite)
```

Tuition and elite correlate, so it is not surprising that graduation rate increases with the elitism of schools.
