---
title: "Week 1 — Exercise 2.9"
author: "Kári Hlynsson"
output: 
  bookdown::html_document2:
    toc: true
    toc_float:
      toc_collapsed: false    
      smooth_scroll: false
    number_sections: false
    toc_depth: 3
---

```{r include = FALSE} 
panel.hist <- function(x, ...)
{
    usr <- par("usr")
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, ...)
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y)
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    text(0.5, 0.5, txt, cex = 1.5)
}

pair_plot <- function(x) { pairs(x, diag.panel = panel.hist, upper.panel = panel.cor) }
```

# Part (a)
Which of the predictors are quantitative, and which are qualitative?

```{r message = FALSE}
library(tidyverse)
library(ISLR2)

data(Auto)
attach(Auto)
```

  
The quantitative variables are `mpg`, `cylinders`, `displacement`, `horsepower`, `weight`, `acceleration`, `year` whereas
the qualitative ones are `name` and `origin`.

# Part (b)
What is the range of each quantitative predictor? You can answer this using the `range` function.

```{r}
ranges <- sapply(Auto[, 1:7], range)
rownames(ranges) <- c("min", "max")

ranges
```

# Part (c)
What is the mean and standard deviation of each quantitative predictor?

```{r}
means <- sapply(Auto[, 1:7], function(c) list(mean = mean(c), sd = sd(c)))
means
```

# Part (d)
Now remove the 10th through 85th observations. What is the range, mean and standard deviation
of each predictor in the subset of the data that remains?

```{r}
ssAuto <- Auto[-c(10:85),]

means <- sapply(Auto[, 1:7], function(c) list(min = min(c), max = max(c), mean = mean(c), sd = sd(c)))
means
```

# Part (e)
Using the full data set, investigate the predictors graphically, using scatterplots or other tools
of your choice. Create some plots highlighting the relationships among the predictors. Comment on your findings.

```{r}
pair_plot(Auto)
```

```{r}
par(mfrow = c(2, 2))

plot_rel <- function(col1, col2) {
  plot(col1, col2, xlab = substitute(col1), ylab = substitute(col2))
  abline(lm(col2 ~ col1), col = "red", lwd = 1.5)
}


plot_rel(weight, displacement)
plot_rel(horsepower, acceleration)
plot_rel(weight, acceleration)
plot_rel(displacement, acceleration)
```

Couple of very strong linear relationships, shown above.

- Displacement is an expression of engine weight, so it is not surprising it correlates with total weight
- Horsepower and acceleration show a decreasing relationship, perhaps because increased horsepower correlated
  with increased engine size and thus more weight.
- Weight and acceleration are negatively correlated, same jutification as above.
- Same as above.

# Part (f)
Suppose that we wish to predict gas mileage (`mpg`) on the basis of the other variables. Do your plots suggest that any of the
other variables might be useful in predicting `mpg`? Justify your answer.

Yes, `displacement`, `weight`, `horsepower`, `acceleration` and `year` all seem to be useful for this task at a first glance.
Let's take a look visually, and add some fits to illustrate the relationship more clearly.

```{r}
par(mfrow = c(2, 2))

plot_quad <- function(col1, col2) {
  plot(col1, col2, xlab = substitute(col1), ylab = substitute(col2))
  lines(lowess(col1, col2), col = 2, lwd = 2)
  
  polfit <- lm(col2 ~ col1 + I(col1^2) + I(col1^2))
  pred <- predict(polfit)
  ord <- order(displacement)
  
  lines(col1[ord], pred[ord], col = 4, lwd = 2)
  
  legend("topright", inset = c(0.025, 0.025), legend = c("LOWESS", "Polynomial"), lwd = 2, col = c(2, 4),
         bty = "n", cex = 0.75)
}

plot_quad(displacement, mpg)
plot_quad(horsepower, mpg)
plot_quad(weight, mpg)
plot_rel(year, mpg)
```

However we might need to take into account interaction effect between variables and collinearity.