---
title: "Exercise 3.10"
author: "Kári Hlynsson"
date: "2024-02-10"
output: 
  bookdown::html_document2:
    toc: true
    toc_float:
      toc_collapsed: false    
      smooth_scroll: false
    number_sections: false
    toc_depth: 3
---

```{r echo = FALSE}
library(equatiomatic)
panel.hist <- function(x, ...)
{
    usr <- par("usr")
    par(usr = c(usr[1:2], 0, 1.5) )
    h <- hist(x, plot = FALSE)
    breaks <- h$breaks; nB <- length(breaks)
    y <- h$counts; y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, ...)
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    par(usr = c(0, 1, 0, 1))
    r <- cor(x, y)
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    text(0.5, 0.5, txt, cex = 1.5)
}

pair_plot <- function(x) { pairs(x, diag.panel = panel.hist, upper.panel = panel.cor) }
```


```{r}
library(ISLR2)
library(MASS)

data(Carseats)
attach(Carseats)

pair_plot(Carseats)
```

# Part (a)
Fit a multiple regression model to predict `Sales` using `Price`, `Urban` and `US`.

```{r}
lin_sales <- lm(Sales ~ Price + Urban + US)
summary(lin_sales)
```

# Part (b)
Provide an interpretation of each coefficient in the model. Be careful -- some of the variables in the model are qualitative.

For quantitative predictors, coefficients can be interpreted as being the slopes in the level set of the predictor $X_j$. For qualitative
predictors this changes to an additive effect if the predictor is in the "positive" state.

# Part (c)
Write out the model in equation form, being careful to handle the qualitative variables properly.

```{r}
extract_eq(lin_sales, use_coefs = TRUE, coef_digits = 4)
```

$$
\operatorname{\widehat{Sales}} = 13.0435 - 0.0545(\operatorname{Price}) - 0.0219(\operatorname{Urban}_{\operatorname{Yes}}) + 1.2006(\operatorname{US}_{\operatorname{Yes}})
$$

# Part (d)
For which of the predictors can you reject the null hypothesis $H_0: \beta_j = 0$?

For all the predictors except `Urban`.


# Part (e)
On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence
of association with the outcome.

```{r}
lin_sales_small <- update(lin_sales, . ~ . - Urban)
summary(lin_sales_small)
```

# Part (f)
How well do the models in (a) and (e) fit the data?

Larger model:

```{r}
par(mfrow = c(2, 2))
plot(lin_sales)
```

Seems to be good residual structure in this model.

Smaller model:

```{r}
par(mfrow = c(2, 2))
plot(lin_sales_small)
```

Similar case here. A couple of observations with high leverage, seemingly only one falls outside of 95% confidence interval
for the studentised residuals and counts as an outlier. These models don't seem to explain very much variation in the data,
$R^2 = 0.2393$ for both models.

We can also compare the models using an $F$ test:

```{r}
anova(lin_sales, lin_sales_small)
```

We fail to reject the null hypothesis that there is a difference between the variation in the data explained by the models,
so naturally we prefer the smaller model `lin_sales_small`.

# Part (g)
Using the model from (e), obtain 95% confidence intervals for the coefficient(s).

```{r}
confint(lin_sales_small)
```

# Part (h)
Is there evidence of outliers or high leverage observations in the model from (e)?

Yes, see my earlier ramblings. 