---
title: "Exercise 7.6"
author: "Kári Hlynsson"
date: "2024-02-13"
output: html_document
---

In this exercise, you will further analyse the `Wage` data set considered throughout
this chapter.

## Part (a)
Perform polynomial regression to predict `wage` using `age`. Use cross-validation to select
the optimal degree $d$ for the polynomial. What degree was chosen, and how does this compare
to the results of hypothesis testing using ANOVA? Make a plot of the resulting polynomial fit
to the data.

Split the data set:

```{r}
library(ISLR2)
library(boot)
library(scales)
data(Wage)

set.seed(1)
```

Fit the polynomials and validate:

```{r}
cv.errs <- sapply(1:10, function(d) {
  glm.fit <- glm(wage ~ poly(age, d), data = Wage)
  cv.err  <- cv.glm(Wage, glm.fit, K = 10)
  return(cv.err$delta[2])
})

plot(cv.errs, type = "b", xlab = "Polynomial degree", ylab = "CV error")
```

The above plot shows the cross-validated test set MSE as a function of the degree of the polynomial fit.
It is indeed consistent with the ANOVA (F-tests) performed in the Lab section of the chapter; we see that
the linear and quadratic models are dominated by the cubic model, and any reduction in the CV error with increasing
polynomial degree is minimal. This would correspond to polynomial models with quartic terms or higher being
statistically insignificant on an ANOVA as seen in the chapter.

A polynomial of degree 3 does the job. Let's look at it graphically:

```{r}
opt.pol  <- lm(wage ~ poly(age, 3), data = Wage)
age.seq  <- seq(18, 80)

opt.pred <- predict(opt.pol, list(age = age.seq), se = TRUE)
opt.se   <- cbind(opt.pred$fit - 1.96 * opt.pred$se.fit,
                  opt.pred$fit + 1.96 * opt.pred$se.fit)

plot(Wage$age, Wage$wage, col = alpha(4, alpha = 0.1), xlab = "Age", ylab = "Wage")
lines(age.seq, opt.pred$fit, lwd = 1.5, col = 4)
matlines(age.seq, opt.se, lwd = 1.25, lty = 2, col = 4)
```

## Part (b)
Fit a step function to predict `wage` using `age`, and perform cross-validation to choose the optimal
number of cuts. Make a plot of the fit observed.

```{r}
cv.errs <- sapply(2:20, function(d) {
  Wage$age.cut <- cut(Wage$age, d)
  
  glm.fit <- glm(wage ~ age.cut, data = Wage)
  cv.err  <- cv.glm(Wage, glm.fit, K = 10)
  return(cv.err$delta[2])
})

plot(2:20, cv.errs, type = "b", xlab = "No. cuts", ylab = "CV error")
```

```{r}
cv.errs
which.min(cv.errs)
```

I believe 8 is best since it lowers the CV error substantially without complexifying the model to an unnecessary degree. We could of course try and see whether this is the case.

```{r}
models <- lapply(2:20, function(d) {
  Wage$age.cut <- cut(Wage$age, d)
  glm.fit <- glm(wage ~ age.cut, data = Wage)
  return(glm.fit)
})

do.call(anova, c(models, test = "F"))
```

Indeed 8 cuts seem to be significant. Make a visual of this step function:

```{r}
step.fit  <- glm(wage ~ cut(age, 8), data = Wage)
step.pred <- predict(step.fit, list(age = age.seq), se = T)
step.se   <- cbind(step.pred$fit - 1.96 * step.pred$se.fit,
                   step.pred$fit + 1.96 * step.pred$se.fit)

plot(wage ~ age, data = Wage, xlab = "Age", ylab = "Wage", col = alpha(4, 0.1))
lines(age.seq, step.pred$fit, lwd = 1.5, col = 4)
matlines(age.seq, step.se, lwd = 1.25, lty = 2, col = 4)
```

