---
title: "Exercise 7.9"
author: "Kári Hlynsson"
date: "2024-02-13"
output: html_document
---

This question uses the variable `dis` (the weighted mean of distances to five Boston employment centres)
and `nox` (nitrogen oxides concentration in parts per 10 million) from the `Boston` data. We will treat `dis`
as the predictor and `nox` as the response.

## Part (a)
Use the `poly()` function to fit a cubic polynomial regression to predict `nox` using `dis`. Report the regression output,
and plot the resulting data and polynomial fits.

```{r}
library(ISLR2)
data(Boston)
attach(Boston)

set.seed(1)

poly.fit <- glm(nox ~ poly(dis, 3), data = Boston)
summary(poly.fit)
```

All terms significant.

```{r}
dis.grid  <- seq(1.1296, 12.1265, length.out = 100)
poly.pred <- predict(poly.fit, list(dis = dis.grid), se = T)
poly.se   <- cbind(poly.pred$fit - 1.96 * poly.pred$se.fit,
                   poly.pred$fit + 1.96 * poly.pred$se.fit)

plot(nox ~ dis, data = Boston, col = alpha(1, 0.10))
lines(dis.grid, poly.pred$fit, lwd = 1.5)
matlines(dis.grid, poly.se, col = 1, lty = 2)
```


## Part (b)
Plot the polynomial fits for a range of different polynomial degrees (say, from 1 to 10), and report the associated residual 
sum of squares.

```{r}
par(mfrow = c(2, 5))

rsss <- sapply(1:10, function(d) {
  poly.fit <- glm(nox ~ poly(dis, d), data = Boston)
  dis.grid  <- seq(1.1296, 12.1265, length.out = 100)
  poly.pl <- predict(poly.fit, list(dis = dis.grid), se = T)
  poly.se   <- cbind(poly.pl$fit - 1.96 * poly.pl$se.fit,
                     poly.pl$fit + 1.96 * poly.pl$se.fit)
  
  plot(nox ~ dis, data = Boston, col = alpha(1, 0.10))
  lines(dis.grid, poly.pl$fit, lwd = 1.5)
  matlines(dis.grid, poly.se, col = 1, lty = 2)
  
  poly.pred <- predict(poly.fit, Boston)
  rss <- sum((poly.pred - nox)^2)
})
```

Residual sums of squares:

```{r}
rsss
```

```{r}
plot(rsss, xlab = "Polynomial degree", ylab = "RSS", type = "b")
```
 
RSS monotonically decreases with increased polynomial degree.

## Part (c)
Perform cross-validation or another approach to select the optimal degree for the polynomial, and explain your results.

```{r}
cv.errs <- sapply(1:10, function(d) {
  glm.fit <- glm(nox ~ poly(dis, d), data = Boston)
  glm.cv  <- cv.glm(Boston, glm.fit, K = 10)
  return(glm.cv$delta[2])
})

plot(cv.errs, type = "b", xlab = "Polynomial degree", ylab = "CV error")
```

```{r}
cv.errs
which.min(cv.errs)
```

Best polynomial is a quartic polynomial.

## Part (d)
Use the `bs()` function to fit a regression spline to predict `nox` using `dis`. Report the output for the fit using four
degrees of freedom. How did you choose the knots? Plot the resulting fit.

Place knots at roughly uniform intervals.

```{r}
library(splines)
knots <- c(3, 6, 9)

spl.fit  <- glm(nox ~ bs(dis, df = 4), data = Boston)
spl.pred <- predict(spl.fit, list(dis = dis.grid), se = T)
spl.se   <- cbind(spl.pred$fit - 1.96 * spl.pred$se.fit,
                  spl.pred$fit + 1.96 * spl.pred$se.fit)

plot(nox ~ dis, data = Boston, col = alpha(1, 0.10))
lines(dis.grid, spl.pred$fit, lwd = 1.5)
matlines(dis.grid, spl.se, lty = 2, col = 1)
```

## Part (e)
Now fit a regression spline for a range of degrees of freedom, and plot the resulting fits and report the resulting RSS.
Describe the results obtained.

```{r}
par(mfrow = c(2, 5))

rsss <- sapply(1:10, function(d) {
  spl.fit  <- glm(nox ~ bs(dis, knots = knots, degree = d), data = Boston)
  dis.grid <- seq(1.1296, 12.1265, length.out = 100)
  spl.pl   <- predict(spl.fit, list(dis = dis.grid), se = T)
  spl.se   <- cbind(spl.pl$fit - 1.96 * spl.pl$se.fit,
                    spl.pl$fit + 1.96 * spl.pl$se.fit)
  
  plot(nox ~ dis, data = Boston, col = alpha(1, 0.10))
  lines(dis.grid, spl.pl$fit, lwd = 1.5)
  matlines(dis.grid, spl.se, col = 1, lty = 2)
  
  spl.pred <- predict(spl.fit, Boston)
  rss <- sum((spl.pred - nox)^2)
})
```

RSSs:

```{r}
rsss
```
```{r}
plot(rsss, type = "b", xlab = "Spline df.", ylab = "RSS")
```

We are seeing the smallest RSS for 10 degrees of freedom as the RSS decreases monotonically.

## Part (f)
Perform cross-validation or another approach in order to select the best degrees of freedom for a regression spline on this data.
Describe your results.

```{r warning = FALSE}
cv.errs <- sapply(1:10, function(d) {
  spl.fit <- glm(nox ~ bs(dis, knots = knots, degree = d), data = Boston)
  spl.cv  <- cv.glm(Boston, spl.fit, K = 10)
  return(spl.cv$delta[2])
})

par(mfrow = c(1, 2))
plot(cv.errs, type = "b", xlab = "Degrees of freedom", ylab = "CV error")
plot(cv.errs[-c(8:10)], type = "b", xlab = "Degrees of freedom", ylab = "CV error")
```

```{r}
cv.errs
which.min(cv.errs)
```
