---
title: "Exercise 7.8"
author: "Kári Hlynsson"
date: "2024-02-13"
output: html_document
---

Fit some of the non-linear models investigated in this chapter to the `Auto` data set.
Is there evidence for non-linear relationships in this data set? Create some informative
plots to justify your answer.

```{r}
library(ISLR2)
library(boot)
library(scales)
data(Auto)
attach(Auto)

set.seed(1)
```

`mpg` as a function of `weight`.

```{r}
cv.errs <- sapply(1:10, function(d) {
  glm.fit <- glm(mpg ~ poly(weight, d), data = Auto)
  glm.cv  <- cv.glm(Auto, glm.fit, K = 10)
  return(glm.cv$delta[2])
})

plot(cv.errs, type = "b", xlab = "Polynomial degree", ylab = "CV error")
```

$d = 2$ seems to do the best job.

```{r}
which.min(cv.errs)
```

Best model visualisation:

```{r}
glm.fit <- glm(mpg ~ poly(weight, 2), data = Auto)
weight.seq <- seq(1613, 5140)

glm.pred <- predict(glm.fit, list(weight = weight.seq), se = T)
glm.se   <- cbind(glm.pred$fit - 1.96 * glm.pred$se.fit,
                  glm.pred$fit + 1.96 * glm.pred$se.fit)

plot(mpg ~ weight, data = Auto, col = alpha(1, 0.15))
lines(weight.seq, glm.pred$fit, lwd = 1.5)
matlines(weight.seq, glm.se, col = 1, lty = 2)
```

This seems to be a pretty good fit. 

```{r}
summary(glm.fit)
```

Let's try some other relationships. 

`acceleration` vs. `mpg` with splines. Let's try having a knot at 10, 15 and 20. Also let's cross-validate over
the degree of the splines.

```{r warning = FALSE}
library(splines)

knots <- c(10, 15, 20)

cv.errs <- sapply(1:7, function(d) {
  spline.fit <- glm(mpg ~ bs(acceleration, knots = knots, degree = d), data = Auto)
  spline.cv  <- cv.glm(Auto, spline.fit, K = 10)
  return(spline.cv$delta[2])
})

plot(cv.errs, xlab = "Spline degree", ylab = "CV error", type = "b")
```

It seems like a linear spline is best here.

```{r}
acceleration.seq <- seq(8.0, 24.8)
spline.fit  <- glm(mpg ~ bs(acceleration, knots = knots, degree = 1), data = Auto)
spline.pred <- predict(spline.fit, list(acceleration = acceleration.seq), se = T)
spline.se   <- cbind(spline.pred$fit - 1.96 * spline.pred$se.fit,
                     spline.pred$fit + 1.96 * spline.pred$se.fit)

plot(mpg ~ acceleration, data = Auto, col = alpha(1, 0.15))
lines(acceleration.seq, spline.pred$fit, lwd = 1.5)
matlines(acceleration.seq, spline.se, col = 1, lty = 2)
```

This is a non-linear relationship. We could also try fitting a smoothing spline:

```{r warning = FALSE}
smooth.fit <- smooth.spline(acceleration, mpg, cv = TRUE)
smooth.fit$df
```
Let's try it out:

```{r}
plot(mpg ~ acceleration, data = Auto, col = alpha(1, 0.15))
lines(smooth.fit, lwd = 1.5)
```

Cooool.