---
title: "Exercise 4.15"
author: "Kári Hlynsson"
date: "2024-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part (a)

Write a function, `Power()`, that prints out the result of raising 2 to the 3rd power. In other words, your function should compute 23 and print out the results. Hint: Recall that `x^a` raises `x` to the power `a`. Use the `print()` function to output the result.

```{r}
Power  <- function() 2^3
```

## Part (b)

Create a new function, `Power2()`, that allows you to pass any two numbers, `x` and `a`, and prints out the value of `x^a`. You can do this by beginning your function with the line [...]

```{r}
Power2 <- function(x, a) x^a
```

## Part (c)

Using the `Power2()` function that you just wrote, compute 103, 817, and 1313.

```{r}
Power2(3, 8)
```

## Part (d)

Now create a new function, `Power3()`, that actually returns the result `x^a` as an R object, rather than simply printing it to the screen. That is, if you store the value `x^a` in an object called result within your function, then you can simply `return()` this result, using the following line:

```{r}
Power3 <- function(x, a) {
  res <- x^a
  return(res)
}
```

## Part (e)

```{r}
x <- 1:10
y <- Power3(x, 2)

ggplot(data.frame(x, y), aes(x, y)) +
  geom_line() +
  labs(
      title = "Plot of f(x) = x²"
  )
```

## Part (f)

```{r}
PlotPower <- function(xs, a) {
  ys <- Power3(xs, a)
  ggplot(
    data.frame(xs, ys),
    aes(xs, ys)
  ) + geom_line()
}

PlotPower(1:10, 3)
```

## 
