---
title: "Exercise 4.16"
author: "Kári Hlynsson"
date: "2024-01-21"
output: html_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(e1071)
library(pROC)
library(class)
library(caret)
library(tidyverse)
library(GGally)
```

Using the Boston data set, fit classification models in order to predict whether a given census tract has a crime rate above or below the median. Explore logistic regression, LDA, naive Bayes, and KNN models using various subsets of the predictors. Describe your findings.

*Hint: You will have to create the response variable yourself, using the variables that are contained in the Boston data set.*

**Solution.** Start by importing the data set.

```{r}
library(ISLR2)
data(Boston)
```

Summary look at the data set.

```{r}
glimpse(Boston)
```

Pair plot:

```{r}
ggpairs(Boston, progress = FALSE)
```

Create the response classifier:

```{r}
Boston <- Boston |> mutate(
  crim01 = ifelse(crim > median(crim), 1, 0)
)
```

